---
title: "multi_arm_msm"
author: "Chang Ho Yoon"
date: "12/01/2024"
output: html_document
---

This markdown details the treatment weight model decisions for emulation of
the multi-arm target trial. As detailed below, the same script may be used
for a high-resistance breakpoint definition for amoxicillin-clavulanate in
E. coli bacteraemia, and also for Enterobacterales bacteraemia by switching
the loaded dataframe or not filtering to E. coli (from the loaded Enterobacterales
dataframe). Some of the code here has been inspired by various papers by
Hernán, Robins et al. 


```{r setup, include=FALSE}
# load dependencies
if (!require("pacman")) install.packages("pacman")
library(pacman)

pacman::p_load(
  tidyverse,    # data management + ggplot2 graphics
  ggthemes,     # expansion pack for ggplot2
  cowplot,      # another expansion pack for ggplot2
  patchwork,    # facet grid
  scales,       # scale functions for ggplot2 visualisation
  stringr,      # manipulate text strings 
  purrr,        # loop over objects in a tidy way
  glue,         # R's version of f-strings
  lubridate,    # time / date objects
  assertr,      # assert functions
  gtsummary,    # publication-qual. tables 
  
  # survival and other regression analyses
  lmtest,       # likelihood-ratio tests
  effectsize,   # utilities to work with indices of effect sizes for variety of models
  survival,     # classic survival analysis package, KM, Cox, etc.
  survminer,    # to draw survival graphs (publication-ready)
  boot,         # bootstrapping functions
  splines,      # fitting splines
  AICcmodavg,    # calculating AIC/AICc for model selection etc.
  survey,       # weighted analyses
  sandwich,     # sandwich estimator
  
  # specific causal inference packages
  ipw,          # IPW calculating package ### NO LONGER MAINTAINED FROM 2021 ###
  WeightIt     # another IPW-calculating package
  )

# create additional custom commands
`%notin%` <- Negate(`%in%`)
`%notwithin%` <- Negate(`%within%`)
`%+%` <- ggplot2::`%+%`

# import functions
source("utils/CAUSAL_utils.R")
```

# import data
```{r}
# enterobacterales data
entero_df = read_csv('#####/#####.csv', 
                     guess_max = 1e6, 
                     show_col_types = FALSE) 

# [OPTIONAL] filter for ecoli_cases
ecoli_df = entero_df %>% filter(`E. coli` == 1)
ecoli_df$id <- 1:nrow(ecoli_df)

ecoli_cases <- ecoli_df %>%
  select(id, ClusterID, SpellID, NewCulturePeriodNumber, ActiveBaseline) %>%
  distinct() 

# outcomes df
outcomes_df = entero_df %>%
  select(ClusterID, SpellID, CensorStatus_day14, OS_14, Followup_outcome_day14, 
         in_hospital_death_day14, out_of_hospital_death_day14)

deaths = ecoli_cases %>% select(id, ClusterID, SpellID) %>%
  left_join(outcomes_df, by=c('ClusterID', 'SpellID')) %>% 
  select(id, LinkedDeathdateAdjusted, LinkedDeathdate_v_adm) %>% distinct()
```


#Load analysis data frames
*set variables*
```{r}
labs_var = c('crp', 'albumin', 'alkaline_phosphatase',
             'urea', 'monocytes', 'neutrophils', 'immature_granulocytes')
labs_var_base = paste0(labs_var, '_base')
obs_var = c('AVPU', 'SupplementaryOxygen', 'OxygenSaturation',
            'HeartRate', 'SBP', 'RespiratoryRate', 'Temperature')
obs_var_base = paste0(obs_var, '_base')
```

*make outcome model df*
```{r}
# set variables
labs_var = c('crp', 'albumin', 'alkaline_phosphatase', 'urea', 'monocytes', 'neutrophils', 'immature_granulocytes')
labs_var_base = paste0(labs_var, '_base')

obs_var = c('AVPU', 'SupplementaryOxygen', 'OxygenSaturation',
            # following obs not in original Cox model
            'HeartRate', 'SBP', 'RespiratoryRate', 'Temperature')
obs_var_base = paste0(obs_var, '_base')

###### EDIT according to user's own location of data ######
# should be truncated already (as per methods)
df_full <- read_csv('#####/#####.csv', guess_max=1e6,
                        show_col_types = FALSE)

# complete cases
df_baseline = df_full %>% filter(time==0)
df_base_comp = df_baseline[complete.cases(df_baseline),]
# ids of complete cases
comp_ids = df_base_comp %>% distinct(id) %>% .$id
df_complete = df_full %>% filter(id %in% comp_ids)

# identify cases receiving active treatment at baseline vs. not
ec.never.id = df_complete %>% group_by(id) %>% filter(all(treat_up == 0)) %>% ungroup() %>% distinct(id) %>% .$id
ec.inactive.base.id = df_complete %>% group_by(id) %>% filter(!any(treat_up == 1 & time == 0)) %>% ungroup() %>% distinct(id) %>% .$id
ec.post.active.id = df_complete %>% group_by(id) %>% filter(any(treat_up == 1 & time > 0) & !any(treat_up == 1 & time == 0)) %>% ungroup() %>% distinct(id) %>% .$id
ec.active.base.id = df_complete %>% group_by(id) %>% filter(any(treat_up == 1 & time == 0)) %>% ungroup() %>% distinct(id) %>% .$id

length(ec.never.id) + length(ec.post.active.id) + length(ec.active.base.id) == length(comp_ids)

# scale variables for improved interpretability
outcome.model.df <- df_complete %>%
  mutate(AgeAtAdmission = AgeAtAdmission / 10,
         BMI = BMI / 10,
         alkaline_phosphatase = (alkaline_phosphatase + 1e-10) / 100,
         alkaline_phosphatase_base = (alkaline_phosphatase_base + 1e-10) / 100,
         crp = (crp + 1e-10) / 10,
         crp_base = (crp_base + 1e-10) / 10) 

head(outcome.model.df, 3)
```

*Define IDs where active baseline, never active, post-baseline active*
```{r}
total_ec = outcome.model.df %>% distinct(id) %>% .$id

new_ec_active = outcome.model.df %>%
  filter(time==0 & treat_up==1) %>% distinct(id) %>% .$id

new_ec_never_active = outcome.model.df %>%
  group_by(id) %>% mutate(keep = case_when(all(treat_up == 0) ~ 1, 
                                           TRUE ~ 0)) %>% 
  ungroup() %>%
  filter(keep == 1) %>% distinct(id) %>% .$id

new_ec_post_active = outcome.model.df %>%
  group_by(id) %>% mutate(keep = case_when(any(treat_up == 1 & time>0) & (treat_up ==0 & time==0) ~ 1, 
                                           TRUE ~ 0)) %>% 
  ungroup() %>%
  filter(keep == 1) %>% distinct(id) %>% .$id

length(new_ec_post_active) + length(new_ec_never_active) + length(new_ec_active) == length(total_ec)
```

*see distributions of outcome*
based on discharge/episode end dates
```{r}
df_complete %>% filter(death == 1) %>%
  group_by(time) %>% summarise(sum = sum(death)) %>%
  ggplot(aes(x=time, y=sum)) + geom_bar(stat = 'identity', fill='steelblue') + labs(x = 'time periods', y='total deaths', title = 'deaths by time [disch dates]') + 
  theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey95"))
```


*check distributions of deaths across time between the two arms*
```{r}
df_complete %>% filter(death == 1) %>%
  group_by(time, treat_base) %>% summarise(sum = sum(death)) %>%
  mutate(active_baseline = case_when(treat_base == 1 ~ 1, TRUE ~ 0)) %>%
  mutate(active_baseline = factor(active_baseline, levels = c(1, 0 ))) %>%
  ggplot(aes(x=time, y=sum, fill=active_baseline)) + 
  geom_bar(stat = 'identity', position = 'stack') + 
  labs(x = 'time periods', y='total deaths', title = 'deaths by time and baseline treatment group [disch dates]') + 
  theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey95"))

```

*unadjusted KM time to death by baseline activity*
```{r}
km_df = outcome.model.df %>%
  left_join(ecoli_cont_outcomes, by='id') %>% distinct() %>%
  group_by(id) %>% slice(1) %>%
  mutate(active_baseline = case_when(treat_base == 1 ~ 1, TRUE ~ 0)) %>%
  mutate(active_baseline = factor(active_baseline, levels = c(1, 0 ))) 

fit <- survfit(Surv(OS_14, CensorStatus_day14) ~ active_baseline, data = km_df)

# plot using custom x-axis time intervals
km_plot <- ggsurvplot(fit, data = km_df, 
           risk.table=TRUE, fontsize =1,
           ylim=c(0.85, 1),
           xlab = "Time intervals",
           break.x.by = 0.25,
           pval = TRUE, pval.coord = c(13, 0.95), pval.size = 3.5, conf.int = FALSE,
           legend.labs = c("Active", "Inactive"),
           legend.title="Baseline antibiotic activity:")

# extract table object from ggsurvplot
tab <- km_plot$table
tab$layers = NULL # clear labels
tab <- tab + 
  geom_text(aes(x = time, y = rev(strata), label = llabels), data = tab$data[tab$data$time %in% c(0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 2.25, 3, 4, 5, 6,7,8,9,10,11,12,13, 14),], size=2) +
  scale_x_continuous(limits = c(0.5, 14), breaks = c(0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 2.25, 3, 4, 5, 6,7,8,9,10,11,12,13, 14),
                     labels = c('0','1','2','3','4','5','6','7','8','9','10','11','12','13','14','15','16','17','18','19')) +
  labs(y = '')

km_ggplot <- km_plot$plot +
  scale_x_continuous(limits = c(0.5, 14), breaks = c(0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 2.25, 3, 4, 5, 6,7,8,9,10,11,12,13, 14),
                     labels = c('0','1','2','3','4','5','6','7','8','9','10','11','12','13','14','15','16','17','18','19'))

km_plot$plot <- km_ggplot
km_plot$table <- tab

km_plot
```



#1. clone
- delay treatment arms 
- arm0 = delay by 0 time intervals
- arm1 = delay by 1 time interval
- arm2 = delay by 2 time intervals
- arm3 = delay by 3 time intervals
- ...
- arm10 = delay by >=10 time intervals


```{r}
# convert tibbles to data table
ec_active_base_df = outcome.model.df %>% filter(id %in% new_ec_active)
ec_inactive_base_df = outcome.model.df %>% filter(id %notin% new_ec_active)

ec_active_base_df = as.data.table(ec_active_base_df, keep.rownames = FALSE)
ec_inactive_base_df = as.data.table(ec_inactive_base_df, keep.rownames = FALSE)

# clone for different strategies
arm0 <- copy(ec_active_base_df)[,arm:=0]
setcolorder(arm0, c('id', 'arm'))

make_arm <- function(df, arm_n){
  
  arm <- copy(df)[,arm:=arm_n]
  setcolorder(arm, c('id', 'arm'))
  
}

list_of_arms = list()
n_arms = 10
for(i in seq_along(1:n_arms)){
  
  output_arm = make_arm(ec_inactive_base_df, arm_n = i)
  
  list_of_arms[[i]] <- output_arm
  
}

arm1 = list_of_arms[[1]]
arm2 = list_of_arms[[2]]
arm3 = list_of_arms[[3]]
arm4 = list_of_arms[[4]]
arm5 = list_of_arms[[5]]
arm6 = list_of_arms[[6]]
arm7 = list_of_arms[[7]]
arm8 = list_of_arms[[8]]
arm9 = list_of_arms[[9]]
arm10 = list_of_arms[[10]]

head(arm10,3)
```


#2. censor
```{r}
### Censor clones in arm 1 when non-adherent ###

# Create a function to define censoring protocol for those assigned to 1-interval delay strategy
arm1.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if ( n> 1 & treat_up[2] == 0) {
    first.cens = 2
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm1.c <- copy(arm1)[,my.censor := arm1.censor(treat_up),by=id][,good := cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm1.c %>%
  group_by(id) %>%
  filter(n()>2) %>% ungroup() %>% distinct(id)

arm1.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)

arm1.c
```

```{r}
### Censor clones in arm 2 when non-adherent ###

# Create a function to define censoring protocol for 2-interval delay
arm2.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 0) {
    first.cens = 3
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm2.c <- copy(arm2)[,my.censor :=arm2.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm2.c %>%
  group_by(id) %>%
  filter(n()>3) %>% ungroup() %>% distinct(id)

arm2.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)

arm2.c
```

```{r}
### Censor clones in arm 3 when non-adherent ###

# Create a function to define censoring protocol for 3-interval delay
arm3.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 0) {
    first.cens = 4
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm3.c <- copy(arm3)[,my.censor :=arm3.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm3.c %>%
  group_by(id) %>%
  filter(n()>4) %>% ungroup() %>% distinct(id)

arm3.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)

arm3.c
```


```{r}
### Censor clones in arm 4 when non-adherent ###

# Create a function to define censoring protocol for 4-interval delay
arm4.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 1) {
    first.cens = 4
  } else if (n > 4 & treat_up[5] == 0) {
    first.cens = 5
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm4.c <- copy(arm4)[,my.censor :=arm4.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm4.c %>%
  group_by(id) %>%
  filter(n()>5) %>% ungroup() %>% distinct(id)

arm4.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)
```


```{r}
### Censor clones in arm 5 when non-adherent ###

# Create a function to define censoring protocol for 5-interval delay
arm5.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 1) {
    first.cens = 4
  } else if (n > 4 & treat_up[5] == 1) {
    first.cens = 5
  } else if (n > 5 & treat_up[6] == 0) {
    first.cens = 6
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm5.c <- copy(arm5)[,my.censor :=arm5.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm5.c %>%
  group_by(id) %>%
  filter(n()>6) %>% ungroup() %>% distinct(id)

arm5.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)

arm5.c
```


```{r}
### Censor clones in arm 6 when non-adherent ###

# Create a function to define censoring protocol for 6-interval delay
arm6.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 1) {
    first.cens = 4
  } else if (n > 4 & treat_up[5] == 1) {
    first.cens = 5
  } else if ( n > 5 & treat_up[6] == 1) {
    first.cens = 6
  } else if (n > 6 & treat_up[7] == 0) {
    first.cens = 7
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm6.c <- copy(arm6)[,my.censor :=arm6.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm6.c %>%
  group_by(id) %>%
  filter(n()>7) %>% ungroup() %>% distinct(id)

arm6.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)

arm6.c
```


```{r}
### Censor clones in arm 7 when non-adherent ###

# Create a function to define censoring protocol for 7-interval delay
arm7.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 1) {
    first.cens = 4
  } else if (n > 4 & treat_up[5] == 1) {
    first.cens = 5
  } else if ( n > 5 & treat_up[6] == 1) {
    first.cens = 6
  } else if ( n > 6 & treat_up[7] == 1) {
    first.cens = 7
  } else if (n > 7 & treat_up[8] == 0) {
    first.cens = 8
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm7.c <- copy(arm7)[,my.censor :=arm7.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm7.c %>%
  group_by(id) %>%
  filter(n()>8) %>% ungroup() %>% distinct(id)

arm7.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)

arm7.c %>%
  select(id, time, death, treat_up, my.censor)
```



```{r}
### Censor clones in arm 8 when non-adherent ###

# Create a function to define censoring protocol for 8-interval delay
arm8.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 1) {
    first.cens = 4
  } else if (n > 4 & treat_up[5] == 1) {
    first.cens = 5
  } else if ( n > 5 & treat_up[6] == 1) {
    first.cens = 6
  } else if ( n > 6 & treat_up[7] == 1) {
    first.cens = 7
  } else if ( n > 7 & treat_up[8] == 1) {
    first.cens = 8
  } else if (n > 8 & treat_up[9] == 0) {
    first.cens = 9
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm8.c <- copy(arm8)[,my.censor :=arm8.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm8.c %>%
  group_by(id) %>%
  filter(n()>9) %>% ungroup() %>% distinct(id)

arm8.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)
```


```{r}
### Censor clones in arm 9 when non-adherent ###

# Create a function to define censoring protocol for 9-interval delay
arm9.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 1) {
    first.cens = 4
  } else if (n > 4 & treat_up[5] == 1) {
    first.cens = 5
  } else if ( n > 5 & treat_up[6] == 1) {
    first.cens = 6
  } else if ( n > 6 & treat_up[7] == 1) {
    first.cens = 7
  } else if ( n > 7 & treat_up[8] == 1) {
    first.cens = 8
  } else if ( n > 8 & treat_up[9] == 1) {
    first.cens = 9
  } else if (n > 9 & treat_up[10] == 0) {
    first.cens = 10
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm9.c <- copy(arm9)[,my.censor :=arm9.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm9.c %>%
  group_by(id) %>%
  filter(n()>10) %>% ungroup() %>% distinct(id)

arm9.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)
```


```{r}
### Censor clones in arm 10 when non-adherent ###

# Create a function to define censoring protocol for 10-interval delay
arm10.censor <- function(treat_up){
  n <- length(treat_up)
  my.censor <- rep(0,n)
  first.cens <- NA
  
  if (n > 1 & treat_up[2] == 1) {
    first.cens = 2
  } else if (n > 2 & treat_up[3] == 1) {
    first.cens = 3
  } else if (n > 3 & treat_up[4] == 1) {
    first.cens = 4
  } else if (n > 4 & treat_up[5] == 1) {
    first.cens = 5
  } else if ( n > 5 & treat_up[6] == 1) {
    first.cens = 6
  } else if ( n > 6 & treat_up[7] == 1) {
    first.cens = 7
  } else if ( n > 7 & treat_up[8] == 1) {
    first.cens = 8
  } else if ( n > 8 & treat_up[9] == 1) {
    first.cens = 9
  } else if ( n > 9 & treat_up[10] == 1) {
    first.cens = 10
  } else if (n > 10 & treat_up[11] == 0) {
    first.cens = 11
  } 

  if ( !(is.na(first.cens) )) {
    my.censor[first.cens:n] <- rep(1,(n-first.cens+1))
  }   
  
  return(my.censor)
}

# Censor clones
arm10.c <- copy(arm10)[,my.censor :=arm10.censor(treat_up),by=id][,good:=cumsum(my.censor),by=id][good <= 1][,good:=NULL]

# "Censor" check
arm10.c %>%
  group_by(id) %>%
  filter(n()>11) %>% ungroup() %>% distinct(id)

arm10.c %>%
  group_by(id) %>%
  mutate(total_n = n()) %>%
  ungroup() %>%
  distinct(id, .keep_all = T) %>%
  count(total_n)

arm10.c
```


*COMBINE ALL CLONED DATAFRAMES*
```{r}
### COMBINE ALL 
active_arm = make_arm(ec_active_base_df, arm_n = 0)
active_arm = active_arm %>% mutate(my.censor = 0)

# Assumed event order within each time interval: death, censor - one precludes the other
all.arms = rbind(active_arm, arm1.c,arm2.c,arm3.c,arm4.c,arm5.c,arm6.c,arm7.c,arm8.c,arm9.c,arm10.c)[,death := ifelse(my.censor == 1 , NA,death)]

# inspect example
all.arms %>%
  filter(time > 0) %>%
  select(id, arm, time, death, treat_up, my.censor) %>%
  filter(arm==2)

# at time == 0 , no included cases should be censored
all.arms %>% filter(time==0) %>% count(death, my.censor)


```

```{r}
# sanity check cases 

## 1: patient starting treatment at 10h has 10 clones
###### - arm1 should have entire follow-up
###### - arms 2-10 should have 1 period of observation

outcome.model.df %>% filter(treat_up == 1) %>% group_by(id) %>% arrange(id, time) %>% slice(1) %>% ungroup() %>%
  filter(time == 1)

all.arms %>% filter(id == 121) %>% select(arm,  id, time, death, treat_up)

## 2: patient starting treatment at 22h has 10 clones
###### - arm3 should have entire follow-up
###### - arm1 censored at time=1, arm2 censored at time=2, arms 4 onwards should be censored at time=3 (when up-treated)

outcome.model.df %>% filter(treat_up == 1) %>% group_by(id) %>% arrange(id, time) %>% slice(1) %>% ungroup() %>%
  filter(time == 3)

all.arms %>% filter(id == 9) %>% select(arm, id, time, death, treat_up)
```



#3. IPTWs using the original 2-arm treatment model to predict artificial censoring in each arm
Ultimately, the aim here is to re-weight at each time interval the patients who are compliant. 

```{r}
# treatment models fit on 2-arm treatment model: active at baseline vs. never active
num_model.formula = as.formula(paste0("treat_up ~ ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10)) + AgeAtAdmission + BMI + immunosuppression + PriorHospitalisation + Specialty + palliative_care_any + firstline_resist + urine_firstline_resist + news_score_base + crp_base + albumin_base + urea_base + neutrophils_base + AVPU_base + OxygenSaturation_base + SupplementaryOxygen_base + alkaline_phosphatase_base + ns(monocytes_base, knots = var_quants$monocytes_base)"))

denom_model.formula = as.formula(paste0("treat_up ~ ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10)) + AgeAtAdmission + BMI + immunosuppression + PriorHospitalisation + Specialty + palliative_care_any + firstline_resist + urine_firstline_resist + worsening_news_class + AVPU + OxygenSaturation + SupplementaryOxygen + albumin + urea + neutrophils + alkaline_phosphatase + ns(monocytes, knots = var_quants$monocytes) + ns(news_score, knots = var_quants$news_score) + crp + last24h_crp_post24_3L"))
```

*interaction model update*
```{r}
denom_model.formula = update(denom_model.formula, .~.
                           + PriorHospitalisation * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + palliative_care_any * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + Specialty * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + firstline_resist * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + urine_firstline_resist  * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + AVPU * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + albumin * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + ns(monocytes, knots = var_quants$monocytes) * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + ns(news_score, knots = var_quants$news_score) * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10))
                           + last24h_crp_post24_3L * ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10)))
```



```{r}
# combined df for all arms up to point of switch
treat.arm0 = outcome.model.df %>% filter(time==0) %>%
  mutate(arm = 0)

treat.all.arms = all.arms %>%
  select(-my.censor) %>%
  filter(arm != 0) %>%
  rbind(treat.arm0) %>%
  mutate(keep = case_when(arm == 0 & time == 0 ~ 1,
                          arm == 1 & time <= 1 ~ 1,
                          arm == 2 & time <=2 ~ 1,
                          arm == 3 & time <=3 ~ 1,
                          arm == 4 & time <=4 ~ 1,
                          arm == 5 & time <=5 ~ 1,
                          arm == 6 & time <=6 ~ 1,
                          arm == 7 & time <=7 ~ 1,
                          arm == 8 & time <=8 ~ 1,
                          arm == 9 & time <=9 ~ 1,
                          arm == 10 & time <=10 ~ 1)) %>%
  filter(keep==1) %>% select(-keep)
```


## [a] Single treatment model for all arms
Estimate probabilities of switching from inactive to active 
- Assumes that the effect of covariates on switching to active is the same
at all timepoints from antibiotic(s) initiation onwards


```{r}
# max follow-up time for post-baseline treatment escalation = when treatment escalation occurred

switch_max_time = outcome.model.df %>%
  filter(id %in% new_ec_post_active) %>%
  filter(treat_up == 1) %>%
  group_by(id) %>%
  arrange(id, time) %>%
  slice(1) %>%
  mutate(max_tstop = tstop) %>%
  dplyr::select(id, max_tstop)

# in treatment model, keep:
# 1) only 1st time row for active baseline
# 2) only up to and incl. the time row when treatment escalation occurs
# 3) all time rows for inactive baseline and not escalated

treat.model.df = outcome.model.df %>%
  left_join(switch_max_time, by='id') %>% distinct() %>%
  filter(id %notin% new_ec_post_active | (id %in% new_ec_post_active & tstop <= max_tstop)) %>%
  dplyr::select(-max_tstop) %>%
  
  mutate(include = case_when(id %in% new_ec_active & time == 0 ~ 1,
                             id %notin% new_ec_active ~ 1,
                             TRUE ~ 0)) %>%
  filter(include == 1) %>% dplyr::select(-include)

# convert tbl --> data.frame for compatibility 
treat.model.df = as.data.frame(treat.model.df)
```


*see distributions of treatment escalation*
```{r, fig.height = 5, fig.width = 8}
n_switch = outcome.model.df %>%
  filter(id %in% new_ec_post_active) %>% distinct(id) %>% nrow(.)

n_inact_base = outcome.model.df %>%
  filter(id %notin% new_ec_active) %>% distinct(id) %>% nrow(.)

t_to_switch = outcome.model.df %>%
  mutate(treat_up = as.numeric(as.character(treat_up))) %>%
  filter(id %in% new_ec_post_active) %>%
  filter(treat_up == 1) %>%
  group_by(id) %>%
  arrange(id, time) %>%
  slice(1) %>%
  mutate(max_tstop = tstart) %>%
  
  group_by(time) %>% summarise(sum = sum(treat_up)) %>%
  rbind(c(15, 0)) %>% rbind(c(16, 0)) %>% rbind(c(17, 0)) %>% rbind(c(18,0)) %>% rbind(c(19,0)) %>%
  ggplot(aes(x=time, y=sum)) + geom_bar(stat = 'identity', fill='steelblue') + 
  labs(x = 'time intervals', y='n episodes with switch to active antibiotics', 
       title = paste0(n_switch, '/',n_inact_base,' switch to active antibiotics after baseline')
       ) + 
  theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey95"))
  
t_to_switch
```


***ONLY UP TO TSTART = 10 DUE TO SPARSITY***
```{r}
end_tstart = 10

new.treat.df = treat.model.df %>%
  filter(tstart <= end_tstart)
```


*obtain knot values for continuous variables*
```{r}
# code to generate splines for continuous variables
other_cont = c('AgeAtAdmission', 'BMI', 'news_score', 'news_score_base')
cont_vars = c(labs_var, obs_var[obs_var %notin% c('AVPU', 'SupplementaryOxygen')],
              labs_var_base, obs_var_base[obs_var_base %notin% c('AVPU_base', 'SupplementaryOxygen_base')], 
              other_cont)

var_quants = list()
for (i in seq_along(cont_vars)){
  var = cont_vars[[i]]
  var_str = as.character(var)

  qs = unname(quantile(new.treat.df[[var]], probs = c(0.1, 0.5, 0.9)))
  
  var_quants[[var_str]] <- qs
  
}
```


```{r}
# treatment model
model_num <- glm(num_model.formula,
                 data = new.treat.df, family = binomial(link = "logit"))

model_denom <- glm(denom_model.formula,
                   data = new.treat.df, family = binomial(link = "logit"))

# compute IPTW
iptw_manual <- as.data.frame(treat.all.arms)
iptw_manual$propensity_num = predict(model_num, iptw_manual, type='response')
iptw_manual$propensity_denom = predict(model_denom, iptw_manual, type='response')

iptw_manual = iptw_manual %>%
  
  # Probability of observed outcome
  mutate(propensity_num_outcome = ifelse(treat_up == 1, propensity_num, 1 - propensity_num),
         propensity_denom_outcome = ifelse(treat_up == 1, propensity_denom, 1 - propensity_denom)) %>% 
  # Numerator / denominator
  mutate(weights_no_time = propensity_num_outcome / propensity_denom_outcome) %>% 
  group_by(id, arm) %>% 
  mutate(ipw_manual = cumprod(weights_no_time)) %>% 
  mutate(ipw_manual = round_half_up(ipw_manual, digits = 6)) %>%
  ungroup() %>%
  mutate(base_treatment = case_when(treat_base == 1 ~ 'active',
                                    treat_base == 0 ~ 'inactive'),
         base_treatment = factor(base_treatment, levels = c('active', 'inactive')))


head(iptw_manual, 3)
summary(iptw_manual$ipw_manual) 
```

*attach IPTWs to corresponding clones*
```{r}
end_tstart = 10

# add weights to person-time data
iptw_link = iptw_manual %>%
  # only cases to time <=10 due to data sparsity beyond
  filter(tstart <= end_tstart) %>%
  dplyr::select(arm, id, time, ipw_manual) 

arm0.outcome = treat.arm0 %>% 
  filter(id %notin% arm0$id) %>%
  rbind(arm0) %>%
  arrange(arm, id, time)

all.arms.outcome <- all.arms %>%
  select(-my.censor) %>%
  filter(arm != 0) %>%
  rbind(arm0.outcome) %>%
  arrange(arm, id, time)
  
all.arms.outcome.iptw <- all.arms.outcome %>%
  left_join(iptw_link, by=c('arm', 'id', 'time')) %>% distinct() %>%
  group_by(arm, id) %>%
  arrange(arm, id, time) %>%
  fill(ipw_manual, .direction = 'down') %>% ungroup() %>%
  assert(not_na, ipw_manual)
```



*fit outcome model*
```{r}
arm_factored = all.arms.outcome.iptw %>% 
  mutate(arm = as.factor(arm))

arm_factored = as.data.frame(arm_factored)

options(warn=-1) # Suppress warning messages

multi.arm.wt.fit <- glm(formula = death==1 ~  arm + ns(time, knots = c(3, 4, 5, 8, 12), Boundary.knots = c(2, 16)),
                
                        family = binomial(link = 'logit'),
                        data = arm_factored,
                        weights = ipw_manual)

# Print results 
summary(multi.arm.wt.fit)

exp(cbind(OR = coef(multi.arm.wt.fit), confint(multi.arm.wt.fit)))

robust_se(multi.arm.wt.fit)
options(warn=-0) 
```


##[b] Separate treat model by time interval
```{r}
options(warn=-1) # Suppress warning messages

time_treat_ipws = list()
# other arms
for(i in seq(0, 10, 1)){
  
  t_df = new.treat.df %>%
    filter(time == i)
  
  t_num <- glm(num_model.formula,
                 data = t_df, family = binomial(link = "logit"))

  if(i <= 4){
    t_denom.formula = update(denom_model.formula, .~. -last24h_crp_post24_3L -last24h_crp_post24_3L*ns(time, knots = c(2, 3, 6, 8), Boundary.knots = c(1, 10)))
  }else{
    t_denom.formula = denom_model.formula
  }
  
  t_denom <- glm(t_denom.formula,
                     data = t_df, family = binomial(link = "logit"))
  
  # compute IPTW
  t.ipw <- as.data.frame(t_df)
  t.ipw$propensity_num = predict(t_num, t.ipw, type='response')
  t.ipw$propensity_denom = predict(t_denom, t.ipw, type='response')
  
  t.ipw = t.ipw %>%
    
    # Probability of observed outcome
    mutate(propensity_num_outcome = ifelse(treat_up == 1, propensity_num, 1 - propensity_num),
           propensity_denom_outcome = ifelse(treat_up == 1, propensity_denom, 1 - propensity_denom)) %>% 
    # Numerator / denominator
    mutate(weights_no_time = propensity_num_outcome / propensity_denom_outcome) %>% 
    mutate(base_treatment = case_when(treat_base == 1 ~ 'active',
                                      treat_base == 0 ~ 'inactive'),
           base_treatment = factor(base_treatment, levels = c('active', 'inactive')))
  
  time_treat_ipws[[i+1]] <- t.ipw
    
}

all.times.ipws = bind_rows(time_treat_ipws) %>%
  verify(nrow(.) == nrow(new.treat.df)) %>%
  assert(not_na, weights_no_time) %>%
  
  group_by(id) %>% 
  mutate(ipw_manual = cumprod(weights_no_time)) %>%
  mutate(ipw_manual = round_half_up(ipw_manual, digits = 6)) %>%
  ungroup()
  
all.times.ipws %>% select(id, ipw_manual) %>% summary(.)

# truncate at 1+99th centiles
threshold_99 <- quantile(all.times.ipws$ipw_manual, 0.99)[[1]]
threshold_1 <- quantile(all.times.ipws$ipw_manual, 0.01)[[1]]
all.times.ipws$sw_1_99 <- all.times.ipws$ipw_manual
# treat_elig$sw_1_99[treat_elig$sw_99 > threshold_99] <- threshold_99
all.times.ipws = all.times.ipws %>% mutate(sw_1_99 = case_when(sw_1_99 > threshold_99 ~ threshold_99,
                                                       sw_1_99 < threshold_1 ~ threshold_1,
                                                       TRUE ~ sw_1_99))

all.times.ipws %>% select(id, ipw_manual, sw_1_99) %>% summary(.)

cat(paste("sd for untruncated ipw:",sd(all.times.ipws$ipw_manual)))
cat(paste("sd for truncated ipw:", sd(all.times.ipws$sw_1_99)))

options(warn=-0) # Suppress warning messages
```

*plot IPTWs across time*
```{r, fig.height = 8, fig.width = 12}
# plot IPWs by time point
iptw_means = compute_iptw_means(all.times.ipws)

p = plot_iptws(all.times.ipws, iptw_means) +
  scale_y_continuous(breaks = c(0, 1, 2, 3, 4, 5), limits = c(0, 5.5) ) +
  labs(
      # title = paste0('[IPTW]: mean = ', mean(all.times.ipws$ipw_manual),
                      # ' sd = ', sd(all.times.ipws$ipw_manual)), 
       x='time intervals', y='IPTW') +
  theme_minimal()

# plot IPWs by antibiotic activity
p2 = ggplot(subset(all.times.ipws, base_treatment == 'active'), 
                       aes(x = propensity_denom, fill = base_treatment)) +
                       geom_histogram(aes(y = ..density..), binwidth=0.01, show.legend=FALSE) + 
                       geom_histogram(data = subset(all.times.ipws, base_treatment == 'inactive'),
                       aes(x = propensity_denom, y = - ..density.., fill = base_treatment), binwidth=0.01, show.legend=FALSE) + 
                       ylab("Density (%)") + 
                       xlab("probability of receiving active treatment") + 
                       # ggtitle("Propensity scores in active and Inactive Baseline\n(Density histogram)") + 
                       scale_fill_discrete(name = "Baseline antibiotic activity") +
  scale_y_continuous(limits = c(-6, 8), breaks = seq(-6, 8, 2)) +
  theme_minimal()

p3 = ggplot(subset(all.times.ipws, base_treatment == 'active'), 
                       aes(x = ipw_manual, fill = base_treatment)) +
                       geom_histogram(aes(y = ..density..), binwidth=0.01) + 
                       geom_histogram(data = subset(all.times.ipws, base_treatment == 'inactive'),
                       aes(x = ipw_manual, y = - ..density.., fill = base_treatment), binwidth=0.01) + 
                       ylab("Density (%)") + 
                       xlab("IPTW") + 
                       # ggtitle("IPTW in Active and Inactive Baseline\n(Density Histogram)") + 
                       scale_fill_discrete(name = "Baseline antibiotic activity") +
  # scale_y_continuous(limits = c(-6, 11), breaks = seq(-6, 10, 2)) +
  theme_minimal()

mean(all.times.ipws$ipw_manual)
sd(all.times.ipws$ipw_manual)

(p2 | p3) / p +
  plot_annotation(tag_levels = 'A')
```

*add IPTWs to each clone*
```{r}
end_tstart = 10

# add weights to person-time data
all.times.iptw_link = all.times.ipws %>%
  # only cases to time <=10 due to data sparsity beyond
  filter(tstart <= end_tstart) %>%
  dplyr::select(id, time, ipw_manual, sw_1_99) %>%
  rename("ipw_by_time" = "ipw_manual",
         "sw_1_99_by_time" = "sw_1_99")

all.arms.outcome.iptw = all.arms.outcome.iptw %>% 
  left_join(all.times.iptw_link, by=c('id', 'time')) %>% distinct() %>% 
  group_by(arm, id) %>%
  arrange(arm, id, time) %>% 
  fill(ipw_by_time, sw_1_99_by_time, .direction = 'down') %>% ungroup() %>%
  assert(not_na, ipw_by_time)

all.arms.outcome.iptw %>%
  select(id, arm, time,  ipw_manual, sw_1_99_by_time, ipw_by_time) %>% 
  summary(.)
```

*using untruncated IPTWs*
```{r}
arm_factored = all.arms.outcome.iptw %>% 
  mutate(arm = as.factor(arm))

arm_factored = as.data.frame(arm_factored)

options(warn=-1) # Suppress warning messages

fit.by.time <- glm(formula = death==1 ~  arm + ns(time, knots = c(3, 4, 5, 8, 12), Boundary.knots = c(2, 16)),
                
                        family = binomial(link = 'logit'),
                        data = arm_factored,
                        weights = ipw_by_time)

# Print results 
summary(fit.by.time)

exp(cbind(OR = coef(fit.by.time), confint(fit.by.time)))

robust_se(fit.by.time)
options(warn=-0) 
```

*using truncated IPTWs*
```{r}
options(warn=-1) # Suppress warning messages

trunc.fit.by.time <- glm(formula = death==1 ~  arm + ns(time, knots = c(3, 4, 5, 8, 12), Boundary.knots = c(2, 16)),
                
                        family = binomial(link = 'logit'),
                        data = arm_factored,
                        weights = sw_1_99_by_time)

# Print results 
summary(trunc.fit.by.time)

exp(cbind(OR = coef(trunc.fit.by.time), confint(trunc.fit.by.time)))

robust_se(trunc.fit.by.time)
options(warn=-0) 
```


*using untruncated with time interaction*
```{r}
options(warn=-1) # Suppress warning messages

fit.by.time <- glm(formula = death==1 ~  arm + ns(time, knots = c(3, 4, 5, 8, 12), Boundary.knots = c(2, 16))
                   + arm*ns(time, knots = c(3, 4, 5, 8, 12), Boundary.knots = c(2, 16)),
                
                        family = binomial(link = 'logit'),
                        data = arm_factored,
                        weights = ipw_by_time)

# Print results 
summary(fit.by.time)

exp(cbind(OR = coef(fit.by.time), confint(fit.by.time)))

robust_se(fit.by.time)
options(warn=-0)
```


*using truncated with time interaction*
```{r}
options(warn=-1) # Suppress warning messages

trunc.fit.by.time <- glm(formula = death==1 ~  arm + ns(time, knots = c(3, 4, 5, 8, 12), Boundary.knots = c(2, 16))
                         + arm*ns(time, knots = c(3, 4, 5, 8, 12), Boundary.knots = c(2, 16)),
                
                        family = binomial(link = 'logit'),
                        data = arm_factored,
                        weights = sw_1_99_by_time)

# Print results 
summary(trunc.fit.by.time)

exp(cbind(OR = coef(trunc.fit.by.time), confint(trunc.fit.by.time)))

robust_se(trunc.fit.by.time)
options(warn=-0) 
```



