---
title: "baseline_models_enterobacterales"
author: "Chang Ho Yoon"
date: "12/01/2024"
output: html_document
---


```{r setup, include=FALSE}
# load dependencies
library(tidyverse)
library(assertr)
library(glue)
library(car)
library(gtsummary)

library(survival)
library(survminer)
library(mfp)
library(MASS)
library(ggeffects)
library(fastDummies)

# create additional custom commands
`%notin%` <- Negate(`%in%`)
`%notwithin%` <- Negate(`%within%`)
`%+%` <- ggplot2::`%+%`

# import functions
source("utils/entero_bsi_modelling_utils.R")
```


*import data*
```{r}
### ALL VARIABLES
categorical_variables <- c('LinkedSex', 'Ethnicity', 'Specialty', 
                           'AdmissionTimeOfDay', 'AdmissionDayOfWeek', 
                           'OutOfHoursAdmission', 'PriorHospitalisation', 
                           'diabetes', 'dialysis', 'palliative', 'immunosuppression', 
                           'SupplementaryOxygen', 'AVPU', 
                           'CommunityOnset', 'PriorEcoliBacteraemiaEpisode', 
                           'Polymicrobial', 
                           'ESBL', 'Ecoli', 'OtherESCAPPM', 'NonEnterobacterales', 
                           'Other', 'Enterobacter', 'Klebsiella',
                           'urinary_source', 'chest_source', 'abdominal_source', 
                           'other_source', 'surgical_source', 'Multi_source', 
                           'PriorAmoxicillin', 'PriorCoamoxiclav', 
                           'PriorAnyAntibiotic', 'PriorBetaLactam', 
                           'InactiveCoamoxiclav', 'InactiveAminoglycoside', 
                           'InactiveOther', 'ActiveAminoglycoside', 'ActiveOther', 
                           'ActiveCoamoxiclav', 'ActiveNonCoamox',
                           'GivenBaselineAminoglycoside', 'GivenMultiAntibioticBaseline', 
                           'GivenTrimethoOrNitro', 
                           'ActivePreCultureAntibiotics', 'ActivePreCultureAminoglycoside'
                           )


continuous_variables <- c('AgeAtAdmission', 'IMDPercentile',
                          'PriorAdmissionsTotalDuration', 
                          'Charlson_sum', 'Elixhauser_sum', 'Height', 'BMI', 
                          'SBP', 'DBP', 'HeartRate', 'RespiratoryRate', 
                          'OxygenSaturation', 'Temperature', 
                          'albumin', 'alkaline_phosphatase', 'alt', 'aptt', 'ast', 
                          'bilirubin', 'calcium_adjusted', 'creatinine', 'egfr', 
                          'eosinophils', 'haematocrit', 'haemoglobin', 
                          'immature_granulocytes', 'lymphocytes', 'magnesium', 
                          'mch', 'mcv', 'monocytes', 'neutrophils', 'phosphate', 
                          'platelets', 'potassium', 'pt', 'sodium', 'urea', 
                          'white_cells'
                          )

# all predictor variables
predictor_variables <- c(categorical_variables, continuous_variables)

# all outcome variables
outcome_variables <- c('OverallSurvival_days_adjusted', 'OverallSurvival_days_capped', 
                       'CensorStatus_day30', 'CensorStatus_day365', 
                       'LengthOfStayFromCulture')
model_variables <- c(predictor_variables, outcome_variables)
```

```{r}
### IMPORT DATA
##### should be modified to read csv of cleaned data from user's directory
# entero_pre_analysis_outcomes_df <- read_csv("/######/######.csv", 
#                                             guess_max = 1e6)

# factorise categorical variables and set references
entero_pre_analysis_outcomes_df <- factorise_categorical(entero_pre_analysis_outcomes_df, 
                                                        factorise_all = TRUE, 
                                                        categorical_variables, 
                                                        factorise_multilevel = TRUE)

# make unfiltered dataframe
unfiltered_survival_df <- make_survival_df(entero_pre_analysis_outcomes_df, 
                                           winsorise_continuous = FALSE)

unfiltered_survival_df
```


###Base effects model

Remove highly correlated variables
```{r}
to_remove <- c('NonEnterobacterales', 'Other', 'Enterobacter', 'Klebsiella', 
               'OtherESCAPPM', 'Polymicrobial', 'OutOfHoursAdmission', 
               'Multi_source', 'AdmissionTimeOfDay',  'creatinine', 'Charlson_sum',
               'haematocrit', 'mch', 'egfr', 'white_cells')

predictor_variables <- predictor_variables[predictor_variables %notin% to_remove]
categorical_variables <- categorical_variables[categorical_variables %notin% to_remove]
continuous_variables <- continuous_variables[continuous_variables %notin% to_remove]

model_variables <- c(predictor_variables, outcome_variables)
```

Winsorise/truncate, scaled and complete dataframes
- scaled for ease of interpretability for selected variables
```{r}
### winsorise / truncate
# symmetrical winsorisation: for continuous covariates with a natural limit
sym_wins_vars <- c('BMI', 'Height', 'Weight', 'SBP', 'DBP', 'HeartRate', 
                   'RespiratoryRate', 'Temperature', 
                   'Charlson_sum', 'Elixhauser_sum', 
                   'potassium', 'sodium', 'albumin', 'creatinine', 
                   'haemoglobin', 'mcv')

# asymmetrical winsorisation: for continuous covariates without a natural limit 
asym_wins_upper_vars <- c('bilirubin', 'alkaline_phosphatase', 'alt', 'urea',
                          'eosinophils', 'lymphocytes', 'immature_granulocytes', 
                          'neutrophils', 'platelets', 'monocytes',
                          'PriorAdmissionsTotalDuration')
asym_wins_lower_vars <- c('OxygenSaturation')

residual_cont_vars <- continuous_variables[continuous_variables %notin% c(sym_wins_vars, asym_wins_lower_vars, asym_wins_upper_vars)]

cat('\n\ncovariates without winsorisation:', paste(residual_cont_vars, collapse = ' + '),'\n')

winsorised_unfiltered_survival_df <- make_survival_df(entero_pre_analysis_outcomes_df, 
                                                        winsorise_continuous = TRUE, 
                                                        sym_wins_vars = sym_wins_vars, 
                                                        sym_limits = c(0.025, 0.975), 
                                                        
                                                        asym_upper_vars = asym_wins_upper_vars, 
                                                        asym_upper_limits = c(0, 0.95),
                                                        
                                                        asym_lower_vars = asym_wins_lower_vars, 
                                                        asym_lower_limits = c(0.05, 1))

winsorised_incomplete_survival_df <- winsorised_unfiltered_survival_df[, model_variables]


### scale survival df
scaled_winsorised_incomplete_survival_df <- winsorised_incomplete_survival_df %>%
  mutate(Height = Height,
         Weight = Weight,
         SBP = SBP/10,
         DBP = DBP/10,
         HeartRate = HeartRate/10,
         RespiratoryRate = RespiratoryRate/10,
         OxygenSaturation = OxygenSaturation,
         Temperature = Temperature,
         albumin = albumin,
         alkaline_phosphatase = alkaline_phosphatase/100,
         alt = alt/10,
         bilirubin = bilirubin/10,
         creatinine = creatinine/10,
         haemoglobin = haemoglobin/10,
         mcv = mcv/10,
         platelets = platelets/100,
         potassium = potassium,
         sodium = sodium,
         urea = urea,
         AgeAtAdmission = AgeAtAdmission/10,
         )

### complete data
scaled_survival_df <- scaled_winsorised_incomplete_survival_df[complete.cases(scaled_winsorised_incomplete_survival_df), ]

head(scaled_survival_df,5)
```


###Univariable estimate for inactive baseline
```{r}
univar_df <- scaled_winsorised_incomplete_survival_df[,c('InactiveBaseline', 'OverallSurvival_days_capped', 'CensorStatus_day30')]
univar_df <- univar_df[complete.cases(univar_df),]
nrow(univar_df)
summary(categorical_univariate_fit('InactiveBaseline', survival_dataframe = univar_df))
```


###Multivariable Cox
```{r}
# choose variables for automatic scaling (i.e., variables that are not selected to be scaled in lines 156-177 above)
# note that scaling does not affect variable selection nor HR estimates - it serves only to improve interpretability
for_auto_scaling <- c('eosinophils', 'immature_granulocytes', 'monocytes', 
                      'lymphocytes', 'neutrophils', 'PriorAdmissionsTotalDuration', 
                      'Charlson_sum', 'Elixhauser_sum')
for_auto_scaling <- for_auto_scaling[for_auto_scaling %in% continuous_variables]
continuous_var_autoscaling <- paste0('fp(', for_auto_scaling,', df = 4, scale=TRUE)')

# choose variables for no scaling:
not_for_scaling <- continuous_variables[continuous_variables %notin% c(for_auto_scaling)]
continuous_var_unscaled <- paste0('fp(', not_for_scaling,', df = 4, scale=FALSE)')

# create formula
if(identical(for_auto_scaling, character(0))){

  multivar_mfp_formula <- as.formula(paste0('Surv(OverallSurvival_days_capped, CensorStatus_day30) ~ (',
                                     paste(categorical_variables, collapse=" + "), ' + ',
                                     paste(continuous_var_unscaled, collapse = " + "), ')'))
}else{
  
  multivar_mfp_formula <- as.formula(paste0('Surv(OverallSurvival_days_capped, CensorStatus_day30) ~ (',
                                     paste(categorical_variables, collapse=" + "), ' + ',
                                     paste(continuous_var_unscaled, collapse = " + "), ' + ',
                                     paste(continuous_var_autoscaling, collapse = " + "), ')'))
}

# fit provisional model
mic.provisional <- mfp(multivar_mfp_formula, 
                     family = cox,
                     data = scaled_survival_df,
                     select = 0.05,           #variate selection significance
                     alpha = 0.05,            #fractional polynomial selection significance
                     method = 'efron',
                     keep = c("InactiveBaseline"),          #keep variable
                     verbose = TRUE,
                     maxits = 20)

# see coefficients
mic.provisional$fit %>%
  tbl_regression(exponentiate = TRUE, conf.int = 0.95, pvalue_fun = ~style_pvalue(.x, digits = 3))
```

Refit onto larger dataset
*notably maintaining the same transforms chosen with the smaller dataset (otherwise at risk of overfitting)*
```{r}
# variables of provisional model
pre_backselection_covar <- c(continuous_variables, categorical_variables)
provisional_base_coef <- names(mic.provisional$coefficients)
provisional_base_covar <- unique(str_extract(provisional_base_coef, paste0(pre_backselection_covar, collapse = '|')))

# complete cases for covariates selected in provisional Cox model:
refit_survival_df <- scaled_winsorised_incomplete_survival_df[, c(provisional_base_covar, 'OverallSurvival_days_capped', 'CensorStatus_day30')]
# sanity check that subset worked for right number of covariates:
length(provisional_base_covar)+2 == ncol(refit_survival_df)

refit_survival_df <- refit_survival_df[complete.cases(refit_survival_df), ]


# add back non-linear transformations as per mic.provisional$fit above
refit_formula <- as.formula(paste0('Surv(OverallSurvival_days_capped, CensorStatus_day30) ~ (',
                            paste(provisional_base_covar[provisional_base_covar %notin% c('monocytes', 'urea')], collapse=" + "), ' + ',
                            'I((monocytes + 0.1)^-0.5) + log(urea)', ')'))

# refit on larger dataset
mic.provisional.refit <- mfp(refit_formula,
                     family = cox,
                     data = refit_survival_df,
                     select = 1,           #variate selection significance
                     alpha = 1,            #fractional polynomial selection significance
                     method = 'efron',
                     keep = c("InactiveBaseline"),          #keep variable
                     verbose = TRUE,
                     maxits = 20)

# see model coefficients
mic.provisional.refit$fit %>%
  tbl_regression(exponentiate = TRUE, conf.level = 0.95, pvalue_fun = ~style_pvalue(.x, digits = 3))
```

Forward selection 
```{r}
output <- add_MFP_term(mic.provisional.refit, 
                       scaled_winsorised_incomplete_survival_df, 
                       categorical_variables, 
                       continuous_variables, 
                       not_for_scaling) 

output[[1]]
```

Base effects model
```{r}
# new complete cases dataframe
base_coef <- names(mic.provisional.refit$coefficients)
base_covar <- unique(str_extract(base_coef, paste0(pre_backselection_covar, collapse = '|')))
base_covar <- c(base_covar, 'neutrophils', 'sodium', 'RespiratoryRate')

base_survival_df <- scaled_winsorised_incomplete_survival_df[, c(base_covar, 'Ecoli', 'OverallSurvival_days_capped', 'CensorStatus_day30')]
# sanity check that subset worked for right number of covariates:
length(base_covar)+2 == ncol(base_survival_df)

base_survival_df <- base_survival_df[complete.cases(base_survival_df), ]

# update formula as indicated 
base_cox_formula <- update(mic.provisional.refit$formula, .~. +neutrophils +I(neutrophils^0.5) +sodium +RespiratoryRate)

# base effects model 
mic.base <- mfp(base_cox_formula, 
                 family = cox,
                 data = base_survival_df,
                 select = 1,           #variate selection significance
                 alpha = 1,            #fractional polynomial selection significance
                 method = 'efron',
                 keep = c("InactiveBaseline"),          #keep variable
                 verbose = FALSE,
                 maxits = 20)

# see base effects model coefficients
mic.base$fit %>%
  tbl_regression(exponentiate = TRUE, conf.level = 0.95, 
                 estimate_fun = purrr::partial(style_ratio, digits = 3),
                 pvalue_fun = ~style_pvalue(.x, digits = 3)) 
```


