---
title: "cluster_sepsis.Rmd"
author: "Chang Ho Yoon"
date: "13/01/2024"
output: html_document
---

```{r setup, include=FALSE}
### load dependencies
# basic
library(tidyverse)
library(gtsummary)
library(assertr)
library(glue)
library(forcats)

# data preprocessing and analysis
library(caret)
library(fastDummies)
library(bestNormalize)
library(modeldata)
library(ppcor)
library(mice)
library(GGally)

# clustering-related
library(cluster)
library(fpc)
library(factoextra)
library(ppclust)
library(NbClust)
library(ConsensusClusterPlus)
library(ggplot2)
library(ggthemes)
library(scales)
library(cowplot)
library(boot)
library(MASS)

# data visualisation
library(RColorBrewer)
library(wordcloud)
library(reshape2)

# create additional custom commands
`%notin%` <- Negate(`%in%`)
`%notwithin%` <- Negate(`%within%`)
`%+%` <- ggplot2::`%+%`

# import functions
source("utils/cluster_utils.R")
source("utils/sepsis_utils.R")
```

*import data*
```{r}
# load cleaned dataframe for patients satisfying sepsis-3 criteria as detailed
# in Seymour et al. (2019), from user's own local directory

# sepsis_pre_analysis_df <- read_csv('#####/#####.csv', guess_max = 1e6)
```

###Select variables as per Seymour et al. paper (2019)
```{r}
### sepsis variables
categorical_variables <- c('LinkedSex')

continuous_variables <- c('AgeAtAdmission', 'albumin', 'alt', 'ast', 'bicarbonate', 
                          'bilirubin', 'urea', 'chloride', 'crp', 'creatinine',
                          'Elixhauser', 'esr', 'GCS', 'glucose', 'HeartRate', 
                          'haemoglobin', 'immature_granulocytes', 'inr', 'lactate', 
                          'OxygenSaturation', 'po2', 'platelets', 'RespiratoryRate', 
                          'sodium', 'SBP', 'Temperature', 'troponin', 'white_cells')

# all predictor variables
sepsis_vars <- c(categorical_variables, continuous_variables)

outcome_variables <- c('OverallSurvival_days_adjusted', 'OverallSurvival_days_capped', 
                       'CensorStatus_day30', 'CensorStatus_day365', 'LengthOfStayFromCulture')

model_variables <- c(sepsis_vars, outcome_variables)

# number of variables:
cat(paste('total number of variables:', length(sepsis_vars)+length(outcome_variables)))
```

Proportions of missingness by variable
```{r}
# from Sepsis et al. (2019) supplement
variables = c('AgeAtAdmission', 'LinkedSex', 'albumin', 'alt', 'ast', 'bicarbonate', 
              'bilirubin', 'urea', 'chloride', 'crp', 'creatinine',
              'Elixhauser', 'esr', 'GCS', 'glucose', 'HeartRate', 'haemoglobin', 
              'immature_granulocytes', 'inr', 'lactate', 'OxygenSaturation', 'po2',
              'platelets', 'RespiratoryRate', 'sodium', 'SBP', 'Temperature', 
              'troponin', 'white_cells')

values = c(0, 0, 12139, 11707, 11714, 6586, 11763, 6962, 6428, 19769, 6647,
           0, 19596, 14271, 6102, 260, 6258, 17303, 11824, 16180, 458, 16161,
           7024, 448, 6074, 262, 1357, 15757, 7102)

sepsis_missingness_df = as.data.frame(variables)
colnames(sepsis_missingness_df) = c("variable")
sepsis_missingness_df$value = values

sepsis_missingness_df$total_n = 20189

sepsis_missingness_df <- sepsis_missingness_df %>%
  mutate(prop = 100*(value/total_n)) %>%
  mutate(dataset = 'sepsis')

# missingness in IORD
iord_n = nrow(sepsis_pre_analysis_df)

iord_missing <- sepsis_pre_analysis_df %>%
  summarise_all(list(~sum(is.na(.)))) %>%
  pivot_longer(everything()) %>%
  rename('variable' = 'name') %>%
  mutate(total_n = iord_n) %>%
  mutate(prop = 100*(value/total_n)) %>%
  filter(variable %in% sepsis_vars) %>%
  mutate(dataset = 'iord')

# plot missingness
missing_plot <- rbind(iord_missing, sepsis_missingness_df)

missing_plot %>%
  arrange(desc(dataset), desc(value)) %>%
  ggplot(aes(x=variable, y=prop, fill=dataset)) +
  geom_bar(position='dodge', stat='identity') +
  aes(x = fct_inorder(variable)) +
  labs(title = 'comparing missingness', 
       x = 'variable',
       y = '% missing') + 
  theme(axis.text.x = element_text(angle = 60, vjust=1, hjust=1))
```

###Prepare clustering dataframe
```{r}
# winsorise/truncate
# symmetrical winsorisation: for continuous covariates with a natural limit
sym_wins_vars <- c('AgeAtAdmission',
                   'SBP', 'HeartRate', 'RespiratoryRate', 'Temperature', 'GCS',
                   'Elixhauser', 'sodium', 'albumin', 'creatinine', 'haemoglobin', 
                   'glucose', 'lactate', 'po2', 'bicarbonate', 'chloride')

# asymmetrical winsorisation: for continuous covariates without a natural limit 
asym_wins_upper_vars <- c('bilirubin', 'alt', 'ast', 'urea', 'crp', 'troponin', 'inr', 
                          'immature_granulocytes', 'white_cells')
asym_wins_lower_vars <- c('OxygenSaturation')

residual_cont_vars <- continuous_variables[continuous_variables %notin% c(sym_wins_vars, asym_wins_lower_vars, asym_wins_upper_vars)]

cat('\n\ncovariates without winsorisation:', paste(residual_cont_vars, collapse = ' + '),'\n')

# make dataframe with truncated variables
unfiltered_cluster_df <- make_survival_df_sepsis(sepsis_pre_analysis_df, winsorise_continuous = TRUE, 
                                                sym_wins_vars = sym_wins_vars, sym_limits = c(0.025, 0.975), 
                                                asym_upper_vars = asym_wins_upper_vars, asym_upper_limits = c(0, 0.95),
                                                asym_lower_vars = asym_wins_lower_vars, asym_lower_limits = c(0.05, 1))

unfiltered_incomplete_cluster_df <- unfiltered_cluster_df[, c('ClusterID', 'SpellID', 'CulturePositive', outcome_variables,  sepsis_vars)]

summary(unfiltered_incomplete_cluster_df)
```

Random forest imputation of missing values
```{r}
set.seed = 12081023
n.features = 29
pre_impute_df <- unfiltered_incomplete_cluster_df %>%
  select(-outcome_variables[outcome_variables %notin% c('CensorStatus_day30')], -ClusterID, -SpellID, -CulturePositive)

# complete data by chained random forest
### ideally this should be performed 10+ times with mean values taken
data.imputed <- missRanger(pre_impute_df, pmm.k=5, num.trees=1000, seed=set.seed)

# create new unique ID
unfiltered_incomplete_cluster_df <- unfiltered_incomplete_cluster_df %>%
  rowid_to_column(., 'id')

data.imputed <- data.imputed %>% select(-CensorStatus_day30)
data.imputed$id <- unfiltered_incomplete_cluster_df$id
data.imputed <- data.imputed[,c(n.features+1,1:n.features)]

# save imputed csv
# write_csv(data.imputed, '#####/imputed_sepsis_29.csv')
```

Yeo-Johnson normalization with standardization + dummy variables/one-hot encoding
```{r, fig.width=22,fig.height=22}
norm_stand_data = yeo_transform(data.imputed, continuous_variables, standardise = TRUE)

# one-hot encode all categorical using fastDummies
one_hot_df <- dummy_cols(norm_stand_data, select_columns = c(categorical_variables), remove_first_dummy = TRUE, remove_selected_columns = TRUE)

# remove outcome and id variables
cluster_data <- one_hot_df
```


###Clustering
```{r}
# Consensus clustering for optimal k clusters
values_only <- data.matrix(cluster_data)
values_only = values_only[, c(1:n.features+1)]

consensus_clusters = ConsensusClusterPlus(values_only,
                                          maxK=10, # max k clusters
                                          reps=1000, # number of sub-samples
                                          pItem=1, # 100% item resampling = columns = variables
                                          pFeature=0.8, # 100% of features to sample = rows = patients
                                          title="hard k-means",
                                          distance="euclidean",
                                          plot = "pdf",   # save plots as PDF
                                          clusterAlg="km"
                                          
                                          )

print('done')  # to remove verbose output
```

Alternatively, optimal k clusters by majority rule
```{r}
nb <- NbClust(values_only,
              distance = "euclidean",
              method = "kmeans",
              min.nc=2, max.nc=10,
              index = "all")

hist(nb$Best.nc[1,], 
     breaks = max(na.omit(nb$Best.nc[1,])),
     main = "majority method of optimal k for k-means",
     xlab = 'k',
     fill = NULL)

# prettify histogram
opt_k = tibble('method' = names(nb$Best.nc[1,]),
               'k' = unname(nb$Best.nc[1,]))

opt_k %>%
  filter(k>0) %>%
  ggplot(aes(k)) + 
  geom_histogram(binwidth=1, fill='steelblue') +
  labs(title = 'majority method of optimal k for k-means', xlab = 'k') +
  scale_x_continuous(breaks = seq(0, 10, 1)) +
  scale_y_continuous(breaks = seq(0, 10, 2)) +
  theme_minimal()
```


###sepsis_29 (Seymour-29 in thesis): Clusters using k=4 
*hard k-means cluster assignment to dataset*
```{r}
sepsis_29_df <- cluster_data %>% select(-id)
custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4", "#C77CFF")
names(custom_colors) <- c("1", "2", "3", "4")

# run k-means clustering algorithm
sepsis.29.km4 <- eclust(sepsis_29_df, "kmeans", k=4, nstart = 25, seed = 100)

factoextra::fviz_cluster(sepsis.29.km4, data = sepsis_29_df, 
  ellipse.type = "convex",
  repel = TRUE) + scale_color_manual(values = custom_colors) + labs(title = 'sepsis emulation 29 variables') + theme_minimal()

sepsis_29_df$cluster <- sepsis.29.km4$cluster

# convert cluster to factor
sepsis_29_df <- sepsis_29_df %>%
  mutate(cluster = as.character(cluster)) %>%
  mutate(cluster = factor(cluster, levels = c('1', '2', '3', '4')))

# write_csv(sepsis_29_df, 'sepsis_vars_km_4.csv')

summary(sepsis_29_df)
```

Repeat with fuzzy C-means algorithm
```{r}
sepsis_29_fuzzy_df <- cluster_data %>% select(-id) 

fcm.29.k4 <- fcm(sepsis_29_fuzzy_df, m=1.1, centers=4, nstart=25, fixcent = TRUE, numseed = 100)
summary(fcm.29.k4)

# plot clusters
fcm.29.k4.plot <- ppclust2(fcm.29.k4, "kmeans")

custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4", "#C77CFF")
names(custom_colors) <- c("1", "2", "3", "4")

factoextra::fviz_cluster(fcm.29.k4.plot, data = values_only, 
  ellipse.type = "convex",
  repel = TRUE) + scale_color_manual(values = custom_colors) + labs(title = 'soft k-means')
```


###sepsis_21 (Seymour-29 minus 8 variables with missingness >15%)
Identify variables with missingness >15% and prepare dataset
```{r}
# highly missing variables
highly_missing = iord_missing %>%
  filter(prop >= 15 & variable != 'GCS') %>%   # GCS is at cut-off of 15% 
  .$variable

# new dataframe
sub_incomplete_df = unfiltered_incomplete_cluster_df %>%
  select(-all_of(highly_missing))

# random forest multiple imputation
set.seed = 12081023
n.features = 29 - length(highly_missing)

pre_impute_df.minus_missing <- sub_incomplete_df %>%
  select(-outcome_variables[outcome_variables %notin% c('CensorStatus_day30')], -ClusterID, -SpellID, -CulturePositive)

# impute missing values using random forest multiple imputation
data.imputed.minus_missing <- missRanger(pre_impute_df.minus_missing, 
                                         pmm.k=5, 
                                         num.trees=1000, 
                                         seed=set.seed)

# create new unique ID
sub_incomplete_df <- sub_incomplete_df %>%
  rowid_to_column(., 'id')

data.imputed.minus_missing <- data.imputed.minus_missing %>% select(-CensorStatus_day30)
data.imputed.minus_missing$id <- sub_incomplete_df$id
data.imputed.minus_missing <- data.imputed.minus_missing[,c(n.features+1,1:n.features)]

# normalisation, standardisation
continuous_variables <- continuous_variables[continuous_variables %notin% highly_missing]

norm_stand_data = yeo_transform(data.imputed, continuous_variables, standardise = TRUE)

# one-hot encode all categorical using fastDummies
one_hot_df <- dummy_cols(norm_stand_data, 
                         select_columns = c(categorical_variables), 
                         remove_first_dummy = TRUE, 
                         remove_selected_columns = TRUE)

# remove outcome and id variables
cluster_data <- one_hot_df 
```

run clustering
```{r}
sepsis_21_df<- cluster_data %>% select(-id)
custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4", "#C77CFF")
names(custom_colors) <- c("1", "2", "3", "4")

sepsis.21.km4 <- eclust(sepsis_21_df, "kmeans", k=4, nstart = 25, seed = 100)

sepsis_21_df$cluster <- sepsis.21.km4$cluster

# convert cluster to factor
sepsis_21_df <- sepsis_21_df %>%
  mutate(cluster = as.character(cluster)) %>%
  mutate(cluster = factor(cluster, levels = c('1', '2', '3', '4')))
```



###sepsis_71 (all 71 variables in IORD data) using k=4

*set variables*
```{r}
### ALL VARIABLES
categorical_variables <- c('LinkedSex', 'Ethnicity', 
                           'SpecialtyGroup', 
                           'AdmissionTimeOfDay', 'AdmissionDayOfWeek', 'OutOfHoursAdmission', 
                           'PriorHospitalisation', 'diabetes', 'dialysis', 'palliative', 
                           'immunosuppression', 'SupplementaryOxygen', 'AVPU',
                           'CommunityOnset', 
                           'urinary_source', 'respiratory_source', 'abdominal_source', 
                           'other_source', 'cns_source', 'skin_soft_tissue_orthopedic_source', 
                           'unspecific_source', 'unknown_source'
                           )

continuous_variables <- c('AgeAtAdmission', 'Height', 'BMI', 'Weight', 'preAdmissionDays', 
                          'IMDPercentile', 'Charlson', 'Elixhauser', 
                          'SBP', 'DBP', 'GCS', 'HeartRate', 'RespiratoryRate', 
                          'OxygenSaturation', 'Temperature', 
                          'albumin', 'alkaline_phosphatase', 'alt', 'aptt', 'ast', 
                          'bicarbonate', 'bilirubin', 'calcium_adjusted', 'chloride', 
                          'creatinine', 'crp', 'eosinophils', 'esr', 'glucose', 
                          'haematocrit', 'haemoglobin', 'immature_granulocytes', 
                          'inr', 'lactate', 'lymphocytes', 'magnesium', 'mch', 
                          'mcv', 'monocytes', 'neutrophils', 'phosphate', 
                          'platelets', 'po2', 'potassium', 'pt', 'sodium', 
                          'troponin', 'urea', 'white_cells'
                          )

# all predictor variables
predictor_variables <- c(categorical_variables, continuous_variables)

outcome_variables <- c('OverallSurvival_days_adjusted', 'OverallSurvival_days_capped', 
                       'CensorStatus_day30', 'CensorStatus_day365', 'LengthOfStayFromCulture')

model_variables <- c(predictor_variables, outcome_variables)
```

Prepare data: truncate & impute
```{r}
##################### Truncation
# symmetrical winsorisation: for continuous covariates with a natural limit
sym_wins_vars <- c('AgeAtAdmission', 'Height', 'BMI', 'Weight', 'IMDPercentile',
                   'SBP', 'DBP', 'HeartRate', 'RespiratoryRate', 'Temperature', 
                  'Elixhauser', 'Charlson',
                   'sodium', 'albumin', 'bicarbonate', 'calcium_adjusted', 
                  'chloride', 'esr', 'glucose', 'haemoglobin', 'haematocrit', 
                  'glucose', 'lactate', 'magnesium', 'mch', 'mcv', 'phosphate', 
                  'po2', 'potassium')

# asymmetrical winsorisation: for continuous covariates without a natural limit 
asym_wins_upper_vars <- c('bilirubin', 'alkaline_phosphatase', 'alt', 'aptt', 
                          'ast',  'creatinine', 'urea', 'crp', 'troponin', 'inr', 
                          'pt', 'lactate', 'eosinophils', 'immature_granulocytes', 
                          'lymphocytes', 'monocytes', 'neutrophils', 'platelets', 
                          'white_cells', 'preAdmissionDays')
asym_wins_lower_vars <- c('OxygenSaturation', 'GCS')

residual_cont_vars <- continuous_variables[continuous_variables %notin% c(sym_wins_vars, asym_wins_lower_vars, asym_wins_upper_vars)]

cat('\n\ncovariates without winsorisation:', paste(residual_cont_vars, collapse = ' + '),'\n')

unfiltered_cluster_df <- make_survival_df_sepsis_plus(sepsis_pre_analysis_df, winsorise_continuous = TRUE, 
                                                      sym_wins_vars = sym_wins_vars, sym_limits = c(0.025, 0.975), 
                                                      asym_upper_vars = asym_wins_upper_vars, asym_upper_limits = c(0, 0.95),
                                                      asym_lower_vars = asym_wins_lower_vars, asym_lower_limits = c(0.05, 1))

unfiltered_incomplete_cluster_df <- unfiltered_cluster_df[, c('ClusterID', 'SpellID', outcome_variables,  predictor_variables)]

##################### Multiple imputation 
set.seed = 12081023

pre_impute_df <- unfiltered_incomplete_cluster_df %>%
  select(-outcome_variables[outcome_variables %notin% c('CensorStatus_day30')], -ClusterID, -SpellID)

n.features = ncol(pre_impute_df[, c(2:ncol(pre_impute_df))])

# complete data by MICE
### ideally this should be performed >10 times with means of values taken
data.imputed.plus <- missRanger(pre_impute_df, pmm.k=5, num.trees=1000, seed=set.seed)

# create new unique ID
unfiltered_incomplete_cluster_df <- unfiltered_incomplete_cluster_df %>%
  rowid_to_column(., 'id')

data.imputed.plus <- data.imputed.plus %>% select(-CensorStatus_day30)
data.imputed.plus$id <- unfiltered_incomplete_cluster_df$id
data.imputed.plus <- data.imputed.plus[,c(n.features+1,1:n.features)]

head(data.imputed.plus, 3)
```

Prepare data: yeo-johnson transformation + standardisation for continuous_variables
- dummy variables for categorical (one-hot encoding)
```{r}
data.imputed.plus.std = yeo_transform(data.imputed.plus, 
                                      continuous_variables, 
                                      standardise = TRUE)

# one-hot encode all categorical using fastDummies
data.imputed.plus.std.onehot <- dummy_cols(data.imputed.plus.std, 
                                           select_columns = c(categorical_variables), 
                                           remove_first_dummy = TRUE, 
                                           remove_selected_columns = TRUE)

# remove outcome and id variables
data.imputed.plus.std.onehot <- data.imputed.plus.std.onehot %>% select(-id)
```


###Optimal number of k for k-means
*consensus clusters approach*
```{r}
values_only <- data.matrix(data.imputed.plus.std.onehot)
colnames(values_only) <- NULL
values_only

consensus_clusters = ConsensusClusterPlus(values_only,
                                          maxK=10, # max k clusters
                                          reps=1000, # number of sub-samples
                                          pItem=1, # 100% item resampling = columns = variables
                                          pFeature=0.8, # 100% of features to sample = rows = patients
                                          title="hard k-means",
                                          distance="euclidean",
                                          plot = "pdf",   # save plots as PDF
                                          clusterAlg="km"
                                          
                                          )

consensus_clusters
print('done') # to remove verbose output
```

*by majority rule*
```{r}
nb <- NbClust(values_only,
              distance = "euclidean",
              method = "kmeans",
              min.nc=2, max.nc=10,
              index = "all")

hist(nb$Best.nc[1,], 
     breaks = max(na.omit(nb$Best.nc[1,])),
     main = "majority method of optimal k for k-means",
     xlab = 'k',
     fill = NULL)

# prettify histogram
opt_k = tibble('method' = names(nb$Best.nc[1,]),
               'k' = unname(nb$Best.nc[1,]))

opt_k %>%
  filter(k>0) %>%
  ggplot(aes(k)) + 
  geom_histogram(binwidth=1, fill='steelblue') +
  labs(title = 'majority method of optimal k for k-means', xlab = 'k') +
  scale_x_continuous(breaks = seq(0, 10, 1)) +
  scale_y_continuous(breaks = seq(0, 10, 2)) +
  theme_minimal()
```

*apply hard k-means cluster assignment to dataset according to optimal k*
```{r}
custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4", "#C77CFF")
names(custom_colors) <- c("1", "2", "3", "4")

sepsis.plus.km4 <- eclust(data.imputed.plus.std.onehot, "kmeans", k=4, nstart = 25, seed = 100)

# visualise clusters
factoextra::fviz_cluster(sepsis.plus.km4, data = data.imputed.plus.std.onehot, 
  ellipse.type = "convex",
  repel = TRUE) + scale_color_manual(values = custom_colors) + labs(title = 'hard k-means')

data.imputed.plus.std.onehot$cluster <- sepsis.plus.km4$cluster

# convert cluster to factor
data.imputed.plus.std.onehot <- data.imputed.plus.std.onehot %>%
  mutate(cluster = as.character(cluster)) %>%
  mutate(cluster = factor(cluster, levels = c('1', '2', '3', '4')))

summary(data.imputed.plus.std.onehot)
```


###sepsis_61 (Seymour variables plus IORD minus 10 highly correlated)
- remove highly correlated variables (|rho| >= 0.8)
```{r}
# highly correlated variables
to_remove = c('Height', 'chloride', 'creatinine', 'haematocrit', 'mch', 'mcv', 'pt', 'white_cells', 'unknown_source_1', 'preAdmissionDays')
n_to_remove = length(to_remove)

# normalise, standardise
data.imputed.plus.std = yeo_transform(data.imputed.plus, 
                                      continuous_variables, 
                                      standardise = TRUE)

# one-hot encode all categorical using fastDummies
data.imputed.plus.std.onehot <- dummy_cols(data.imputed.plus.std, 
                                           select_columns = c(categorical_variables), 
                                           remove_first_dummy = TRUE, 
                                           remove_selected_columns = TRUE)

# remove outcome and id variables
data.imputed.plus.std.onehot <- data.imputed.plus.std.onehot %>% select(-id)

# remove highly correlated variables
data.imputed.plus.minus <- data.imputed.plus.std.onehot %>% select(-all_of(to_remove))
```

Run clustering
```{r}
custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4", "#C77CFF")
names(custom_colors) <- c("1", "2", "3", "4")

sepsis.plus.minus.km4 <- eclust(data.imputed.plus.minus, "kmeans", k=4, nstart = 25, seed = 100)

factoextra::fviz_cluster(sepsis.plus.minus.km4, data = data.imputed.plus.minus,
  ellipse.type = "convex",
  repel = TRUE)  +
  labs(title = 'hard k-means') +
  scale_color_manual(values = custom_colors)

data.imputed.plus.minus$cluster <- sepsis.plus.minus.km4$cluster

# convert cluster to factor
data.imputed.plus.minus <- data.imputed.plus.minus %>%
  mutate(cluster = as.character(cluster)) %>%
  mutate(cluster = factor(cluster, levels = c('1', '2', '3', '4')))
```



###sepsis_29 + extra continuous variables only

Set variables
```{r}
# set continuous variables 
continuous_variables <- c(
                          'AgeAtAdmission', 'BMI', 'Weight',  'IMDPercentile', 
                          'Charlson', 'Elixhauser', 
                          'SBP', 'DBP', 'GCS', 'HeartRate', 'RespiratoryRate', 
                          'OxygenSaturation', 'Temperature', 
                          'albumin', 'alkaline_phosphatase', 'alt', 'aptt', 'ast', 
                          'bicarbonate', 'bilirubin', 'calcium_adjusted', 'crp',  
                          'eosinophils', 'esr', 'glucose', 'haemoglobin', 
                          'immature_granulocytes', 'inr', 'lactate', 'lymphocytes', 
                          'magnesium',  'monocytes', 'neutrophils', 'phosphate', 
                          'platelets', 'po2', 'potassium', 'sodium', 'troponin', 
                          'urea'
                          )

example_cont_seymour_plus = data.imputed.plus.minus[, c(continuous_variables, "LinkedSex_M")]
```

Run clustering
```{r}
# run clustering
custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4", "#C77CFF")
names(custom_colors) <- c("1", "2", "3", "4")

seymour.plus.minus.cont.km4 <- eclust(example_cont_seymour_plus, 
                                      "kmeans", 
                                      k=4, 
                                      nstart = 25, 
                                      seed = 100)

factoextra::fviz_cluster(seymour.plus.minus.cont.km4, data = example_cont_seymour_plus, 
  ellipse.type = "convex",
  repel = TRUE)  + 
  labs(title = 'hard k-means') + 
  scale_color_manual(values = custom_colors)

example_cont_seymour_plus$cluster <- seymour.plus.minus.cont.km4$cluster

# convert cluster to factor
example_cont_seymour_plus <- example_cont_seymour_plus %>%
  mutate(cluster = as.character(cluster)) %>%
  mutate(cluster = factor(cluster, levels = c('1', '2', '3', '4')))
```

