---
title: "cluster_enterobacterales.Rmd"
author: "Chang Ho Yoon"
date: "12/01/2024"
output: html_document
---


```{r setup, include=FALSE}
### load dependencies
# basic
library(tidyverse)
library(gtsummary)
library(assertr)
library(glue)
library(forcats)

# data preprocessing and analysis
library(caret)
library(fastDummies)
library(bestNormalize)
library(modeldata)
library(ppcor)
library(mice)
library(GGally)

# clustering-related
library(cluster)
library(fpc)
library(factoextra)
library(ppclust)
library(NbClust)
library(ConsensusClusterPlus)
library(ggplot2)
library(ggthemes)
library(scales)
library(cowplot)
library(boot)
library(MASS)

# data visualisation
library(RColorBrewer)
library(wordcloud)
library(reshape2)

# create additional custom commands
`%notin%` <- Negate(`%in%`)
`%notwithin%` <- Negate(`%within%`)
`%+%` <- ggplot2::`%+%`

# import functions
source("utils/cluster_utils.R")
```

*import data*
```{r}
### ALL VARIABLES
categorical_variables <- c('LinkedSex', 'Ethnicity', 
                           'Specialty', 'AdmissionTimeOfDay', 'AdmissionDayOfWeek', 
                           'OutOfHoursAdmission', 'PriorHospitalisation', 
                           'diabetes', 'dialysis', 'palliative', 'immunosuppression', 
                           'SupplementaryOxygen', 'AVPU', 
                           'CommunityOnset', 'PriorBacteraemiaEpisode', 'Polymicrobial', 
                           'ESBL', 'Ecoli', 'OtherESCAPPM', 'NonEnterobacterales', 
                           'Other', 'Enterobacter', 'Klebsiella',
                           'urinary_source', 'chest_source', 'abdominal_source', 
                           'other_source', 'surgical_source', 'Multi_source', 
                           'PriorAmoxicillin', 'PriorCoamoxiclav', 
                           'PriorAnyAntibiotic', 'PriorBetaLactam', 
                           # 'InactiveBaseline', 'InactiveCoamoxiclav', 
                           # 'ActiveAminoglycoside', 'ActiveOther', 
                           # 'ActiveCoamoxiclav', 'ActiveNonCoamox',
                           'GivenBaselineAminoglycoside', 'GivenMultiAntibioticBaseline', 
                           # 'ActivePreCultureAntibiotics', 'ActivePreCultureAminoglycoside', 
                           'GivenTrimethoOrNitro'
                           )

continuous_variables <- c('AgeAtAdmission', 'PriorAdmissionsTotalDuration', 
                          'IMDPercentile', 'Charlson_sum', 'Elixhauser_sum', 
                          'Height', 'BMI', 
                          'SBP', 'DBP', 'HeartRate', 'RespiratoryRate', 
                          'OxygenSaturation', 'Temperature', 
                          'albumin', 'alkaline_phosphatase', 'alt', 'aptt', 'ast', 
                          'bilirubin', 'calcium_adjusted', 'creatinine', 'egfr', 
                          'eosinophils', 'haematocrit', 'haemoglobin', 
                          'immature_granulocytes', 'lymphocytes', 'magnesium', 
                          'mch', 'mcv', 'monocytes', 'neutrophils', 'phosphate', 
                          'platelets', 'potassium', 'pt', 'sodium', 'urea', 
                          'white_cells', 'crp'
                          )

# remove highly correlated (|R|>0.8) variables
to_remove <- c('NonEnterobacterales', 'Other', 'Enterobacter', 'Klebsiella', 
               'OtherESCAPPM', 'Polymicrobial', 'OutOfHoursAdmission', 
               'Multi_source', 'AdmissionTimeOfDay',
               'haematocrit', 'mch', 'egfr', 'white_cells')
categorical_variables <- categorical_variables[categorical_variables %notin% to_remove]
continuous_variables <- continuous_variables[continuous_variables %notin% to_remove]

# all predictor variables
predictor_variables <- c(categorical_variables, continuous_variables)

outcome_variables <- c('OverallSurvival_days_adjusted', 'OverallSurvival_days_capped', 
                       'CensorStatus_day30', 'CensorStatus_day365', 'LengthOfStayFromCulture')

cluster_variables <- c(predictor_variables, outcome_variables)
```

```{r}
### ***EDIT*** load csv of unscaled, winsorised dataframe (all cases, not just complete ones) from user's local directory
### this should be the same csv used for Enterobacterales baseline modelling
# unfiltered_cluster_df <- read_csv("#####/#####.csv", 
#                                   guess_max = 1e6)

unfiltered_incomplete_cluster_df <- unfiltered_cluster_df[, cluster_variables]

# create new unique ID
unfiltered_incomplete_cluster_df <- unfiltered_incomplete_cluster_df %>%
  rowid_to_column(., 'id')

entero_cluster_df <- entero_cluster_df %>%
  rowid_to_column(., 'id')

# scale survival df
scaled_incomplete_cluster_df <- unfiltered_incomplete_cluster_df %>%
  mutate(across(c(AgeAtAdmission, alt, bilirubin, haemoglobin, mcv, SBP, DBP, 
                  HeartRate, RespiratoryRate), ~ ./10 )) %>%
  mutate(across(c(alkaline_phosphatase, creatinine), ~ . /100)) 

head(scaled_incomplete_cluster_df,5)
```


Create clustering dataframe
- complete cases only 
- one-hot encoding / dummy variables for categorical variables
- Yeo-Johnson transformation for continuous variables
```{r}
# complete cases only
cluster_df <- scaled_incomplete_cluster_df[, c('id', cluster_variables)]
complete_cluster_df <- cluster_df[complete.cases(cluster_df),]
complete_ids = complete_cluster_df %>% .$id
complete_cluster_df <- complete_cluster_df %>% select(-id)

# one-hot encode all categorical using fastDummies
one_hot_cluster_df <- dummy_cols(complete_cluster_df, 
                                 select_columns = c(categorical_variables), 
                                 remove_first_dummy = TRUE, 
                                 remove_selected_columns = TRUE)

# yeo_johnson transform continuous variables
all_dims_transformed_df <- yeo_transform(one_hot_cluster_df, 
                                         continuous_variables[continuous_variables %in% cluster_variables])

summary(all_dims_transformed_df)
```


###Optimal number of k for k-means
```{r}
nb <- NbClust(all_dims_transformed_df,
              distance = "euclidean",
              method = "kmeans",
              min.nc=2, max.nc=10,
              index = "all")

hist(nb$Best.nc[1,], 
     breaks = max(na.omit(nb$Best.nc[1,])),
     main = "majority method of optimal k for k-means",
     xlab = 'k',
     fill = NULL)

# prettify histogram
opt_k = tibble('method' = names(nb$Best.nc[1,]),
               'k' = unname(nb$Best.nc[1,]))

opt_k %>%
  filter(k>0) %>%
  ggplot(aes(k)) + 
  geom_histogram(binwidth=1, fill='steelblue') +
  labs(title = 'majority method of optimal k for k-means', xlab = 'k') +
  scale_x_continuous(breaks = seq(0, 6, 1)) +
  theme_minimal()
```

Alternative: optimal number of k for k-means by CONSENSUS CLUSTERING
```{r}
values_only <- data.matrix(all_dims_transformed_df)
colnames(values_only) <- NULL
values_only

consensus_clusters = ConsensusClusterPlus(values_only,
                                          maxK=10, # max k clusters
                                          reps=1000, # number of sub-samples
                                          pItem=1, # 100% item resampling = columns = variables
                                          pFeature=0.8, # 100% of features to sample = rows = patients
                                          title="hard k-means",
                                          distance="euclidean",
                                          plot = "pdf",   # save plots as PDF
                                          clusterAlg="km"
                                          
                                          )

consensus_clusters

print('done') # to clear verbose output
```


###FCM clustering, k=3
```{r}
set.seed(100)
custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4")
names(custom_colors) <- c("1", "2", "3")

res.fcm.k3 <- fcm(all_dims_transformed_df, m=1.1, centers=3, nstart=10, fixcent = TRUE, numseed=100)
res.fcm.k3

summary(res.fcm.k3)

# plot clusters
res.fcm.k3.plot <- ppclust2(res.fcm.k3, "kmeans")

factoextra::fviz_cluster(res.fcm.k3.plot, data = all_dims_transformed_df, 
  ellipse.type = "convex",
  geom="point",
  # palette = "jco",
  repel = TRUE) + scale_color_manual(values = custom_colors) + labs(title = 'soft k-means') + theme_minimal()

```


```{r}
# update data with clusters
soft_clusters <- as.data.frame(res.fcm.k3$u) %>%
  rename('cluster_1' = 'Cluster 1',
         'cluster_2' = 'Cluster 2',
         'cluster_3' = 'Cluster 3')

new_cox_df <- scaled_incomplete_cluster_df[, c('InactiveBaseline', cluster_variables, outcome_variables)]
new_cox_df <- new_cox_df[complete.cases(new_cox_df),]

# add cluster assignments and probabilities to scaled dataframe
soft_complete_cluster_df <- new_cox_df %>%
  cbind(soft_clusters) %>%
  mutate(cluster_probability = pmax(cluster_1, cluster_2, cluster_3)) %>%   # take max probability as the cluster
  mutate(cluster = case_when(cluster_probability == cluster_1 ~ '1',
                             cluster_probability == cluster_2 ~ '2',
                             cluster_probability == cluster_3 ~ '3')) %>%
  mutate(cluster = factor(cluster, levels = c('1', '2', '3')))

# plot cluster probabilities by cluster
custom_colors_pastel <- c("#F4CCCC", "#d9ead3", "#d0e0e3")
names(custom_colors_pastel) <- c("1", "2", "3")

soft_complete_cluster_df %>%
  ggplot(aes(x=cluster, y=cluster_probability, fill=cluster)) +
  geom_boxplot(show.legend=FALSE) +
  geom_jitter(size = 0.4, alpha=0.5, show.legend=FALSE) +
  scale_fill_manual(values = custom_colors_pastel) +
  scale_y_continuous(breaks=seq(0.2, 1.0, 0.1)) + 
  labs(title = 'soft k-means: \ndistribution of cluster assignment probability by cluster', x='cluster', y='cluster probability') +
  theme_minimal()
```


Plot differences between the clusters
```{r}
compare_clusters_df.k3 <- soft_complete_cluster_df %>%
  select(cluster, cluster_probability) %>%
  cbind(cluster_eda) %>% select(cluster_variables)

# one-hot encode all categorical using fastDummies
one_hot_compare_clusters.k3 <- dummy_cols(compare_clusters_df.k3, select_columns = c(categorical_variables), remove_first_dummy = FALSE, remove_selected_columns = TRUE)

# yeo_johnson transform continuous variables
compare_clusters_df_transformed.k3 <- yeo_transform(one_hot_compare_clusters.k3, continuous_variables[continuous_variables %in% model_variables])

summary(compare_clusters_df_transformed.k3)
```

- compare differences for continuous variables
```{r}
cluster_plot_df.k3 <- compare_clusters_df_transformed.k3
cluster_plot_df.k3$cluster <- soft_complete_cluster_df$cluster

# hard-code variable order for plot (to maximise clarity)
variate_order = c('Age', 'Neutrophils', 'Monocytes', 'Immature granulocytes', 
                  'Platelets', 'Urea', 'Creatinine','Potassium', 'CRP', 
                  'Lymphocytes', 'Haemoglobin', 'MCH', 'MCV', 'Eosinophils',  
                  'IMD percentile', 'Respiratory rate', 'Oxygen saturation', 
                  'Heart rate', 'Charlson', 'Elixhauser', 
                  'Prior admissions total days',  'Alkaline phosphatase', 
                  'Sodium', 'BMI', 'Weight', 'ALT', 'Bilirubin',  'SBP', 'DBP', 
                  'Temperature', 'Albumin')

# compute differences in continuous variables
hard_cont_plots.k3 <- compare_continuous_clusters(standardised_df =  cluster_plot_df.k3, 
                                              continuous_var = c(continuous_variables), 
                                              type = 'hard k-means', 
                                              n_clusters = 3, 
                                              variate_order = variate_order)

# plot comparisons between clusters for continuous variable medians
hard_cont_plots.k3[[5]] %>%
  mutate(variate = str_to_title(variate)) %>% 
  mutate(variate = case_when(variate == 'Sbp' ~ 'SBP',
                             variate == 'Dbp' ~ 'DBP',
                             variate == 'Alt' ~ 'ALT',
                             variate == 'Bmi' ~ 'BMI',
                             variate == 'Heartrate' ~ 'Heart rate',
                             variate == 'Mcv' ~ 'MCV',
                             variate == 'Oxygensaturation' ~ 'Oxygen saturation',
                             variate == 'Respiratoryrate' ~ 'Respiratory rate',
                             variate == 'Imdpercentile' ~ 'IMD percentile',
                             variate == 'Alkaline_phosphatase' ~ 'Alkaline phosphatase',
                             variate == 'Ageatadmission' ~ 'Age',
                             variate == 'Prioradmissionstotalduration' ~ 'Prior admissions total days',
                             variate == 'Charlson_sum' ~ 'Charlson',
                             variate == 'Elixhauser_sum' ~ 'Elixhauser',
                             variate == 'Crp' ~ 'CRP',
                             variate == 'Immature_granulocytes' ~ 'Immature granulocytes',
                             TRUE ~ variate)) %>%
  ggplot(mapping = aes(x=variate, y=diff, group=cluster, color=cluster)) +
      geom_line() +
      geom_hline(yintercept = 0, linetype='dashed') +
      aes(x=factor(variate, levels=variate_order)) +
      labs(title = paste0('Enterobacterales k=3: differences of median'),
           x = 'Variate',
           y = 'Difference between standardised cluster median and overall median') +
      theme(panel.background = element_blank(), 
            panel.grid.major = element_line(colour = "grey95"), 
            axis.text.x = element_text(angle = 45, vjust=1, hjust=1))
```

- compare differences for categorical variables
```{r}
# manual ordering for maximal clarity
cat_variate_order = c('Prior hospitalisation',  "Specialty: medical subspecialty",   
                      'Diabetes',  'Sex: male', "Respiratory source", 
                      "Other source", "Surgical source" , 'Supplementary oxygen',  
                      'Immunosuppression', 'ESBL',  "Specialty: other", 'Dialysis',
                      "Admitted Monday", "Admitted Tuesday", "Admitted Wednesday", 
                      "Admitted Thursday", "Admitted Friday", "Admitted Saturday", 
                      "Admitted Sunday", 'Ethnicity: unrecorded', 'Palliative', 
                      'AVPU: pain / unresponsive', 'AVPU: verbal',  'AVPU: alert',    
                      "Abdominal source", "Specialty: Acute general surgery", 
                      'E. coli',   'Community onset', 'Ethnicity: white', 
                      "Specialty: Acute general medicine",  'Urinary source')

cat_comp.k3.sans_prior_abx <- compare_categorical_clusters(cluster_plot_df.k3, 
                                            c(continuous_variables, 'Height', 
                                              "PriorAnyAntibiotic_1",  "PriorCoamoxiclav_1", 
                                              'PriorBetaLactam_1', "PriorAmoxicillin_1"), 
                                            type_clustering = 'hard k-means', 
                                            n_clusters = 3, 
                                            variate_order = cat_variate_order)

cat_comp.k3.sans_prior_abx[[2]] %>%
  mutate(variate = as.character(variate)) %>%
  mutate(variate = case_when(variate == 'PriorHospitalisation_1' ~ 'Prior hospitalisation',
                             variate == 'Specialty_Medical subspecialty' ~ 'Specialty: medical subspecialty',
                             variate == 'diabetes_1' ~ 'Diabetes',
                             variate == 'LinkedSex_M' ~ 'Sex: male',
                             variate == 'chest_source_1' ~ 'Respiratory source',
                             variate == 'other_source_1' ~ 'Other source',
                             variate == 'surgical_source_1' ~ 'Surgical source',
                             variate == 'immunosuppression_1' ~ 'Immunosuppression',
                             variate == 'ESBL_1' ~ 'ESBL',
                             variate == 'Specialty_Other' ~ 'Specialty: other',
                             variate == 'dialysis_1' ~ 'Dialysis',
                             variate == 'SupplementaryOxygen_1' ~ 'Supplementary oxygen',
                             variate == 'AdmissionDayOfWeek_Monday' ~ 'Admitted Monday',
                             variate == 'AdmissionDayOfWeek_Tuesday' ~ 'Admitted Tuesday',
                             variate == 'AdmissionDayOfWeek_Wednesday' ~ 'Admitted Wednesday',
                             variate == 'AdmissionDayOfWeek_Thursday' ~ 'Admitted Thursday',
                             variate == 'AdmissionDayOfWeek_Friday' ~ 'Admitted Friday',
                             variate == 'AdmissionDayOfWeek_Saturday' ~ 'Admitted Saturday',
                             variate == 'AdmissionDayOfWeek_Sunday' ~ 'Admitted Sunday',
                             variate == 'Ethnicity_unrecorded' ~ 'Ethnicity: unrecorded',
                             variate == 'palliative_1' ~ 'Palliative',
                             variate == 'AVPU_Pain / Unresponsive' ~ 'AVPU: pain / unresponsive',
                             variate == 'AVPU_Verbal' ~ 'AVPU: verbal',
                             variate == 'AVPU_Alert' ~ 'AVPU: alert',
                             variate == 'abdominal_source_1' ~ 'Abdominal source',
                             variate == 'Specialty_Acute and general surgery' ~ 'Specialty: Acute general surgery',
                             variate == 'Ecoli_1' ~ 'E. coli',
                             variate == 'CommunityOnset_1' ~ 'Community onset',
                             variate == 'Ethnicity_white' ~ 'Ethnicity: white',
                             variate == 'Specialty_Acute and general medicine' ~ 'Specialty: Acute general medicine',
                             variate == 'urinary_source_1' ~ 'Urinary source',
                             TRUE ~ variate)) %>%
  filter(variate %in% cat_variate_order_sans_prior_abx) %>%
  ggplot(mapping = aes(x=variate, y=diff, group=cluster, color=cluster)) + 
      geom_line() +
      aes(x=factor(variate, levels = cat_variate_order_sans_prior_abx)) +
      labs(title = paste0('Enterobacterales k=3: differences of cluster proportions'), 
           x = 'Variate',
           y = 'Difference of cluster proportions to overall proportions') + 
      theme(panel.background = element_blank(), 
            panel.grid.major = element_line(colour = "grey95"), 
            axis.text.x = element_text(angle = 60, vjust=1, hjust=1))
```


###Word clouds of clusters
```{r}
# to remove prior antibiotoic exposure
prior_abx_vars = c('PriorAnyAntibiotic', 'PriorCoamoxiclav', 'PriorBetaLactam', 
                   'PriorAmoxicillin', 'InactiveBaseline', 'GivenTrimethoOrNitro', 
                   'ActiveOther', 'ActiveAminoglycoside')
```

Prepare dataframe for word clouds
```{r}
# create df suitable for word clouds
# one-hot encode all categorical using fastDummies
cloud_one_hot.k3 <- dummy_cols(compare_clusters_df.k3, 
                               select_columns = c(categorical_variables), 
                               remove_first_dummy = TRUE, 
                               remove_selected_columns = TRUE)

# remove outcome and id variables
cloud_clusters_pre_transform.k3 <- cloud_one_hot.k3[, -which(names(cloud_one_hot.k3) %in% c(outcome_variables, "ClusterID", "NewCulturePeriodNumber"))]

# yeo_johnson transform continuous variables
cloud_clusters_df_transformed.k3 <- yeo_transform(cloud_clusters_pre_transform.k3, continuous_variables[continuous_variables %in% model_variables])

cloud_cluster_df.k3 <- cloud_clusters_df_transformed.k3 %>%
# add hospital onset - inverse of community onset
  mutate(HospitalOnset = case_when(CommunityOnset_1 == 1 ~ 0,
                                   CommunityOnset_1 == 0 ~ 1))

# add cluster assignment from FCM above
cloud_cluster_df.k3$cluster <- soft_complete_cluster_df$cluster

# reconfigure columns to allow word clouds (by frequency / magnitude)
cloud_cat.k3 = prep_categorical_wordcloud(cloud_cluster_df.k3,
                                       cont_vars = c(continuous_variables, 'Height'),
                                       n_clusters = 3)

head(cloud_cat.k3, 5)
```

Plot word clouds
```{r, fig.height=10, fig.width=10}
set.seed(100)

plot_wordcloud(hard_cont_plots.k3[[5]], cloud_cat.k3,
               select_cluster = 1,
               remove = prior_abx_vars)

plot_wordcloud(hard_cont_plots.k3[[5]], cloud_cat.k3,
               select_cluster = 2,
               remove = prior_abx_vars)

plot_wordcloud(hard_cont_plots.k3[[5]], cloud_cat.k3,
               select_cluster = 3,
               remove = prior_abx_vars)
```



###FCM clustering, k=4
Run FCM algorithm
```{r}
set.seed(100)
custom_colors <- c("#F8766D", "#7DAE01", "#00BFC4", "#DB6AFE")
names(custom_colors) <- c("1", "2", "3", "4")

res.fcm.k4 <- fcm(all_dims_transformed_df, 
                  m=1.1, 
                  centers=4, 
                  nstart=10, 
                  fixcent = TRUE, 
                  numseed=100)

summary(res.fcm.k4)

# plot clusters
res.fcm.k4.plot <- ppclust2(res.fcm.k4, "kmeans")

factoextra::fviz_cluster(res.fcm.k4.plot, 
                         data = all_dims_transformed_df, 
                         ellipse.type = "convex",
                         geom='point',
                         # palette = "jco",
                         repel = FALSE) + 
  scale_color_manual(values = custom_colors) + 
  labs(title = 'soft k-means') + 
  theme_minimal()

```

Save cluster assignments and probabilities
```{r}
# update data with clusters
soft_clusters.k4 <- as.data.frame(res.fcm.k4$u) %>%
  rename('cluster_1' = 'Cluster 1',
         'cluster_2' = 'Cluster 2',
         'cluster_3' = 'Cluster 3',
         'cluster_4' = 'Cluster 4')

new_cox_df.k4 <- scaled_incomplete_cluster_df[, c('InactiveBaseline', cluster_variables, outcome_variables)]
new_cox_df.k4 <- new_cox_df.k4[complete.cases(new_cox_df.k4),]

soft_complete_cluster_df.k4 <- new_cox_df.k4 %>%
  cbind(soft_clusters.k4) %>%
  mutate(cluster_probability = pmax(cluster_1, cluster_2, cluster_3, cluster_4)) %>%   # take max probability as the cluster
  mutate(cluster = case_when(cluster_probability == cluster_1 ~ '1',
                             cluster_probability == cluster_2 ~ '2',
                             cluster_probability == cluster_3 ~ '3',
                             cluster_probability == cluster_4 ~ '4')) %>%
  # match groups to k=3 assignments
  mutate(cluster = case_when(cluster == '2' ~ '4',
                             cluster == '4' ~ '3',
                             cluster == '3' ~ '2',
                             TRUE ~ cluster)) %>%
  mutate(cluster = factor(cluster, levels = c('1', '2', '3', '4')))

# add unique id and save assignments
# soft_complete_cluster_df.k4$id = 1:nrow(soft_complete_cluster_df.k4)
# to_save = soft_complete_cluster_df.k4 %>% select(id, cluster, cluster_probability, cluster_1, cluster_2, cluster_3, cluster_4)
# 
# write_csv(to_save, 'entero_k4_clusters.csv')
```

Plot cluster probabilities
```{r}
custom_colors_pastel <- c("#F4CCCC", "#d9ead3", "#d0e0e3", "#EBC3FF")
names(custom_colors_pastel) <- c("1", "2", "3", "4")

soft_complete_cluster_df.k4 %>%
  ggplot(aes(x=cluster, y=cluster_probability, fill=cluster)) +
  geom_boxplot(show.legend=FALSE) +
  geom_jitter(size = 0.4, alpha=0.5, show.legend=FALSE) +
  scale_fill_manual(values = custom_colors_pastel) +
  scale_y_continuous(breaks=seq(0.4, 1.0, 0.1)) + 
  labs(title = 'soft k-means: \ndistribution of cluster assignment probability by cluster', x='cluster', y='cluster probability') +
  theme_minimal()
```

