---
title: "compare_clusters"
author: "Chang Ho Yoon"
date: "13/01/2024"
output: html_document
---

```{r setup, include=FALSE}
# load dependencies
library(cvms)
library(tidyverse)
```


*load data*
```{r}
### each will need 
# Seymour-29, k-means k = 4
seymour_29_df = read_csv('seymour_vars_km_4.csv', guess_max=1e6)

# Seymour-21: Seymour-29 minus highly missing i.e. Seymour-21, k-means k=4 
seymour_21_df = read_csv('seymour_minus_missing_km_4.csv', guess_max=1e6)

# Seymour-29 complete cases only
seymour_29_complete_df = read_csv('seymour_complete.csv', guess_max=1e6)

# Seymour-71: Seymour=29 plus all IORD variables
seymour_71_df = read_csv('seymour_plus_km_4.csv', guess_max=1e6)

# Seymour-61: Seymour-29 plus IORD variables minus highly-correlated
seymour_61_df = read_csv('seymour_plus_minus_km_4.csv', guess_max=1e6)

# Seymour-29 + continuous only: 
seymour_61_plus_cont_df = read_csv('seymour_plus_minus_cont_km_4.csv', guess_max = 1e6)

# Seymour-61 with both k-means k=4 and k-means k=5 clusters
seymour_61_km4_5_df = read_csv('seymour_plus_minus_km_4_and_5.csv', guess_max = 1e6)

### External Seymour model ASSIGNMENT OF PHENOTYPES
ext_phenotypes_df = read_csv('seymour_to_iord_both_assigned.csv', guess_max = 1e6) 
  # %>% select(-`...1`)
```

*seymour-29 vs. seymour-21 (seymour minus missing)*
i.e., >15% missing = bicarbonate, chloride, lactate, troponin, glucose, esr, ast, po2
and RF imputation of the rest
```{r}
cat(paste0('number of variables in Seymour only: ', length(colnames(seymour_29_df))-1))
cat(paste0('\nnumber of variables in Seymour minus highly missing: ', length(colnames(seymour_21_df))-1))

# confusion matrix
confusion = seymour_29_df %>% select(cluster)

confusion$plus <- seymour_21_df$cluster

cm <- as.data.frame(table(confusion$cluster, confusion$plus))

cm <- cm %>% rename('Seymour_only' = 'Var1', 'Seymour_minus_missing' = 'Var2')


plot_confusion_matrix(cm, 
                      target_col = "Seymour_only", 
                      prediction_col = "Seymour_minus_missing",
                      counts_col = "Freq",
                      add_normalized = FALSE,
                      add_counts = TRUE,
                      add_col_percentages = TRUE,
                      add_row_percentages = TRUE,
                      place_x_axis_above = FALSE) +
  labs(title = 'confusion plot: Seymour variables only vs. Seymour minus highly missing', x = 'Seymour 29 variables', y = 'Seymour minus highly missing')
```


*seymour-29 vs. seymour-21 complete cases only*
i.e., >15% missing = bicarbonate, chloride, lactate, troponin, glucose, esr, ast, po2
removed
```{r}
complete_ids = seymour_29_complete_df %>% .$id

cat(paste0('number of variables in Seymour only: ', length(colnames(seymour_29_df))-1))
cat(paste0('\nnumber of variables in Seymour complete cases: ', length(colnames(seymour_29_complete_df))-1))

# confusion matrix
confusion = seymour_29_df[complete_ids,]  %>% select(cluster)

confusion$plus <- seymour_29_complete_df$cluster

cm <- as.data.frame(table(confusion$cluster, confusion$plus))

cm <- cm %>% rename('Complete' = 'Var1', 'Imputed' = 'Var2')

plot_confusion_matrix(cm, 
                      target_col = "Imputed", 
                      prediction_col = "Complete",
                      counts_col = "Freq",
                      add_normalized = FALSE,
                      add_counts = TRUE,
                      add_col_percentages = TRUE,
                      add_row_percentages = TRUE,
                      place_x_axis_above = FALSE) +
  labs(title = 'confusion plot: Seymour imputed vs. Seymour complete', x = 'Seymour with imputation', y = 'Seymour complete cases')
```


*seymour-29 vs. seymour-71*
```{r}
cat(paste0('number of variables in Seymour only: ', length(colnames(seymour_29_df))-1))
cat(paste0('\nnumber of variables, incl. one-hot-encoded, in Seymour plus: ', length(colnames(seymour_71_df))-1))

# confusion matrix
confusion = seymour_29_df %>% select(cluster)

confusion$plus <- seymour_71_df$cluster

cm <- as.data.frame(table(confusion$cluster, confusion$plus))

cm <- cm %>% rename('Seymour_only' = 'Var1', 'Seymour_plus' = 'Var2')


plot_confusion_matrix(cm, 
                      target_col = "Seymour_only", 
                      prediction_col = "Seymour_plus",
                      counts_col = "Freq",
                      add_normalized = FALSE,
                      add_counts = TRUE,
                      add_col_percentages = TRUE,
                      add_row_percentages = TRUE,
                      place_x_axis_above = FALSE) +
  labs(title = 'confusion plot: Seymour variables only vs. Seymour plus', x = 'Seymour 29 variables', y = 'Seymour plus all IORD variables')

```


*seymour-29 vs. seymour-61*
```{r}
cat(paste0('number of variables in Seymour only: ', length(colnames(seymour_29_df))-1))
cat(paste0('\nnumber of variables, incl. one-hot-encoded, in Seymour plus minus: ', length(colnames(seymour_61_df))-1))

# confusion matrix
confusion = seymour_29_df %>% select(cluster)

confusion$plus <- seymour_61_df$cluster

cm <- as.data.frame(table(confusion$cluster, confusion$plus))

cm <- cm %>% rename('Seymour_only' = 'Var1', 'Seymour_plus_minus' = 'Var2')


plot_confusion_matrix(cm, 
                      target_col = "Seymour_only", 
                      prediction_col = "Seymour_plus_minus",
                      counts_col = "Freq",
                      add_normalized = FALSE,
                      add_counts = TRUE,
                      add_col_percentages = TRUE,
                      add_row_percentages = TRUE,
                      place_x_axis_above = FALSE) +
  labs(title = 'confusion plot: Seymour variables only vs. Seymour plus minus highly-correlated', x = 'Seymour 29 variables', y = 'Seymour plus all IORD variables excl. highly-correlated')

cat(paste('rand index:', rand.index(confusion$cluster, confusion$plus)))
cat(paste('\nadjusted rand index:', adj.rand.index(confusion$cluster, confusion$plus)))
```


*seymour-71 vs. seymour-61*
```{r}
cat(paste0('\nnumber of variables, incl. one-hot-encoded, in Seymour plus: ', length(colnames(seymour_71_df))-1))
cat(paste0('\nnumber of variables, incl. one-hot-encoded, in Seymour plus minus: ', length(colnames(seymour_61_df))-1))

# confusion matrix
confusion = seymour_71_df %>% select(cluster)

confusion$plus <- seymour_61_df$cluster

cm <- as.data.frame(table(confusion$cluster, confusion$plus))

cm <- cm %>% rename('Seymour_plus' = 'Var1', 'Seymour_plus_minus' = 'Var2')


plot_confusion_matrix(cm, 
                      target_col = "Seymour_plus", 
                      prediction_col = "Seymour_plus_minus",
                      counts_col = "Freq",
                      add_normalized = FALSE,
                      add_counts = TRUE,
                      add_col_percentages = TRUE,
                      add_row_percentages = TRUE,
                      place_x_axis_above = FALSE) +
  labs(title = 'confusion plot:\nSeymour plus vs. Seymour plus minus highly-correlated', x = 'Seymour plus', y = 'Seymour plus minus')

```


*seymour-61 k = 4 vs. k = 5 (CDF delta plot suggests k = 5 may be optimal)*
```{r}
cat(paste0('\nnumber of variables, incl. one-hot-encoded, in Seymour plus minus km=5: ', length(colnames(seymour_61_km4_5_df))-1))

# confusion matrix
confusion = seymour_61_km4_5_df %>% select(cluster)

confusion$plus <- seymour_61_km4_5_df$cluster_k5

cm <- as.data.frame(table(confusion$cluster, confusion$plus))

cm <- cm %>% rename('Seymour_plus_minus_k_4' = 'Var1', 'Seymour_plus_minus_k_5' = 'Var2')

ggplot(cm, aes(Seymour_plus_minus_k_4, Seymour_plus_minus_k_5)) +                        
  geom_tile(aes(fill = Freq)) +
  scale_fill_gradient(low = "#FFFFFF",
                      high = "#012345") +
  labs(x = 'all IORD variables excl. highly-correlated, k=4', y = 'all IORD variables excl. highly-correlated, k=5')


# plot_confusion_matrix(cm, 
#                       target_col = "Seymour_plus_minus_k_4", 
#                       prediction_col = "Seymour_plus_minus_k_6",
#                       counts_col = "Freq",
#                       add_normalized = FALSE,
#                       add_counts = TRUE,
#                       add_col_percentages = TRUE,
#                       add_row_percentages = TRUE,
#                       place_x_axis_above = FALSE) +
#   labs(title = 'confusion plot:\nSeymour plus minus k=4 vs. Seymour plus minus k=6', x = 'k=4', y = 'k=6')
```



*seymour-29 plus continuous variables only k = 4 (Consensus optimal)*
```{r}
cat(paste0('number of variables in Seymour only: ', length(colnames(seymour_29_df))-1))
cat(paste0('\nnumber of variables, incl. one-hot-encoded, in Seymour plus minus continuous only: ', length(colnames(seymour_61_plus_cont_df))-1))

# confusion matrix
confusion = seymour_29_df %>% select(cluster)

confusion$plus <- seymour_61_plus_cont_df$cluster

cm <- as.data.frame(table(confusion$cluster, confusion$plus))

cm <- cm %>% rename('Seymour_only' = 'Var1', 'Seymour_plus_minus_continuous' = 'Var2')


plot_confusion_matrix(cm, 
                      target_col = "Seymour_only", 
                      prediction_col = "Seymour_plus_minus_continuous",
                      counts_col = "Freq",
                      add_normalized = FALSE,
                      add_counts = TRUE,
                      add_col_percentages = TRUE,
                      add_row_percentages = TRUE,
                      place_x_axis_above = FALSE) +
  labs(title = 'confusion plot:\nSeymour only vs. Seymour plus minus continuous', x = 'Seymour 29 variables', y = 'Seymour plus continuous variables in IORD')

```



###EXTERNAL SEYMOUR ET AL. MODEL APPLIES TO IORD DATA
*Seymour code-assigned clusters vs. IORD consensus clusters*
```{r}
ext_phenotypes_df

# confusion matrix
confusion = ext_phenotypes_df %>% select(cluster)

confusion$iord <- ext_phenotypes_df$iord_cluster

cm <- as.data.frame(table(confusion$cluster, confusion$iord))

cm <- cm %>% rename('Seymour' = 'Var1', 'IORD' = 'Var2')


plot_confusion_matrix(cm, 
                      target_col = "Seymour", 
                      prediction_col = "IORD",
                      counts_col = "Freq",
                      add_normalized = FALSE,
                      add_counts = TRUE,
                      # intensity_by = "normalized",
                      add_col_percentages = TRUE,
                      add_row_percentages = TRUE,
                      place_x_axis_above = FALSE) +
  labs(title = 'Figure 2: confusion plot of\n               internal phenotypes vs.\n               external model phenotypes', x = 'SENECA-derived clusters', y = 'Internal IORD clusters') +
  scale_x_discrete(labels= c('alpha', 'beta', 'gamma', 'delta'))

cat(paste('rand index:', rand.index(ext_phenotypes_df$cluster, ext_phenotypes_df$iord_cluster)))
cat(paste('\nadjusted rand index:', adj.rand.index(ext_phenotypes_df$cluster, ext_phenotypes_df$iord_cluster)))
```

